### 2017-07-29 README.Rmd
add to README. add tests for multiple level treatment

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`drtmle`

[![Travis-CI Build Status](https://travis-ci.org/benkeser/drtmle.svg?branch=master)](https://travis-ci.org/benkeser/drtmle)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/benkeser/drtmle?branch=master&svg=true)](https://ci.appveyor.com/project/benkeser/drtmle)
[![Coverage Status](https://img.shields.io/codecov/c/github/benkeser/drtmle/master.svg)](https://codecov.io/github/benkeser/drtmle?branch=master)
[![CRAN](http://www.r-pkg.org/badges/version/drtmle)](http://www.r-pkg.org/pkg/drtmle)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)

> Nonparametric estimators of the average treatment effect with doubly-robust confidence intervals and hypothesis tests

__Author:__ [David Benkeser](https://www.benkeserstatistics.com/)

---

## Description

`drtmle` is an R package that computes marginal effect estimators for binary 
treatments on continuous and binary outcomes. The targeted minimum loss-based (TMLE) estimators are doubly robust, not only with respect to consistency, but also with respect to asymptotic normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017 (_accepted Biometrika_; [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)). Therefore, it is possible to construct doubly-robust confidence intervals and hypothesis tests. Also included are methods for computing valid confidence intervals for an inverse probability of treatment weighted (IPTW) estimator of the average treatment effect when the propensity score is estimated via super learning, as discussed in [van der Laan, 2014](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).

---

## Installation

- Install the most recent _stable release_:
  `devtools::install_github("benkeser/drtmle")`

- To contribute, install the _development version_:
  `devtools::install_github("benkeser/drtmle", ref = "develop")`

---

## Use

### Doubly-robust inference for average treatment effect

This package can be used to estimate covariate-adjusted marginal means under multiple discrete levels of a treatment. It may be used in situations where data consist of a vector of baseline covariates (`W`), a multi-level treatment assignment (`A`), and a continuous or binary-valued outcome (`Y`). The function `drtmle` may be used to estimate $E[E(Y | A=a_0, W)]$ for user-selected values of $a_0$ (via option `a_0`). The resulting targeted minimum loss-based estimates are doubly robust with respect to both consistency and asymptotic normality. The function computes doubly robust variance estimators that can be used to construct doubly robust confidence intervals for covariate-adjusted marginal means and contrasts between these means (via the `drconfint` function) for different levels of $a_0$. A simple example on simulated data is shown below.

```{r}
# load packages
library(drtmle)
library(SuperLearner)

# simulate simple data structure
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n,-2,2), W2 = rbinom(n,1,0.5))
A <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W2))
Y <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W2 + A))

# estimate the covariate-adjusted marginal mean for A = 1 and A = 0
# here, we do not properly estimate the propensity score
fit1 <- drtmle(W = W, A = A, Y = Y, # input data
              a_0 = c(0,1), # return estimates for A = 0 and A = 1
              SL_Q = "SL.npreg", # use kernel regression for E(Y | A=a, W)
              glm_g = "W1 + W2", # use misspecified main terms glm for E(A | W)
              SL_Qr = "SL.npreg", # use kernel regression to guard against
              					  # misspecification of outcome regression
              SL_gr = "SL.npreg" # use kernel regression to guard against 
              					 # misspecification of propensity score
              )
# print the output
fit1

# get confidence intervals for marginal means
ci_fit1 <- confint(fit1)
# print the output
ci_fit1

# get confidence intervals for ate
ci_ate1 <- confint(fit1,contrast = c(-1,1))
# print the output
ci_ate1
```

### Inference for super learner-based IPTW

The package additionally includes a function for computing valid confidence intervals about an inverse probability of treatment weight (IPTW) estimator when super learning is used to estimate the propensity score. We call this estimator the inverse super learning probability of treatment estimator (ISLPTW).

```{r}
# fit iptw 
fit2 <- islptw(Y = Y, A = A, W = W, a_0 = c(0,1),
               SL_g = c("SL.glm","SL.mean","SL.step.interaction"),
               SL_Qr = "SL.npreg")
# print the output
fit2

# compute a confidence interval for margin means
ci_fit2 <- confint(fit2)
# print the output
ci_fit2

# compute a confidence interval for the ate
ci_ate2 <- confint(fit2, contrast = c(-1,1))
# print the output
ci_ate2
```
---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/benkeser/survtmle/issues).

---

## Citation

After using the `drtmle` R package, please cite the following:

        @article{benkeser2017improved,
          year  = {2017},
          author = {Benkeser, David C and Carone, Marco and van der Laan, Mark J and Gilbert, Peter B},
          title = {Doubly-robust nonparametric inference on the average treatment effect},
          journal = {Biometrika}
        }

---

## License

&copy; 2016-2017 [David C. Benkeser](http://www.benkeserstatistics.com)

The contents of this repository are distributed under the MIT license. See
below for details:
```
The MIT License (MIT)

Copyright (c) 2016-2017 David C. Benkeser

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```