### 2018-05-08 README.Rmd
update readmes

---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`drtmle`

[![Travis-CI Build Status](https://travis-ci.org/benkeser/drtmle.svg?branch=master)](https://travis-ci.org/benkeser/drtmle)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/benkeser/drtmle?branch=master&svg=true)](https://ci.appveyor.com/project/benkeser/drtmle)
[![Coverage Status](https://img.shields.io/codecov/c/github/benkeser/drtmle/master.svg)](https://codecov.io/github/benkeser/drtmle?branch=master)
[![CRAN](http://www.r-pkg.org/badges/version/drtmle)](http://www.r-pkg.org/pkg/drtmle)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/drtmle)](https://CRAN.R-project.org/package=drtmle)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)
[![DOI](https://zenodo.org/badge/75324341.svg)](https://zenodo.org/badge/latestdoi/75324341)

> Nonparametric estimators of the average treatment effect with doubly-robust confidence intervals and hypothesis tests

__Author:__ [David Benkeser](https://www.benkeserstatistics.com/)

---

## Description

`drtmle` is an R package that computes marginal means of an outcome under fixed
levels of a treatment. The package computes targeted minimum loss-based (TMLE)
estimators that are doubly robust, not only with respect to consistency, but
also with respect to asymptotic normality, as discussed in [Benkeser, et al (2017)](https://academic.oup.com/biomet/article-abstract/104/4/863/4554445). This property
facilitates construction of doubly-robust confidence intervals and hypothesis
tests.

The package additionally includes methods for computing valid confidence
intervals for an inverse probability of treatment weighted (IPTW) estimator of
the average treatment effect when the propensity score is estimated via super
learning, as discussed in
[van der Laan (2014)](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).

---

## Installation

Install the current stable release from
[CRAN](https://cran.r-project.org/) via

```{r cran-installation, eval = FALSE}
install.packages("drtmle")
```

A developmental release may be installed from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with:

```{r gh-installation, eval = FALSE}
devtools::install_github("benkeser/drtmle")
```

---

## Usage

### Doubly-robust inference for the average treatment effect

Suppose the data consist of a vector of baseline covariates (`W`), a multi-level
treatment assignment (`A`), and a continuous or binary-valued outcome (`Y`). The
function `drtmle` may be used to estimate $E[E(Y \mid A = a_0, W)]$ for
user-selected values of $a_0$ (via option `a_0`). The resulting targeted minimum
loss-based estimates are doubly robust with respect to both consistency and
asymptotic normality. The function computes doubly robust covariance estimates
that can be used to construct doubly robust confidence intervals for marginal
means and contrasts between means. A simple example on simulated data is shown
below. We refer users to [the vignette](https://benkeser.github.io/drtmle/articles/using_drtmle.html) for more information and further examples.

```{r}
# load packages
library(drtmle)
library(SuperLearner)

# simulate simple data structure
set.seed(12345)
n <- 200
W <- data.frame(W1 = runif(n,-2,2), W2 = rbinom(n,1,0.5))
A <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W1*W$W2))
Y <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W1*W$W2 + A))

# estimate the covariate-adjusted marginal mean for A = 1 and A = 0
# here, we do not properly estimate the propensity score
fit1 <- drtmle(W = W, A = A, Y = Y, # input data
               a_0 = c(0, 1), # return estimates for A = 0 and A = 1
               SL_Q = "SL.npreg", # use kernel regression for E(Y | A = a, W)
               glm_g = "W1 + W2", # use misspecified main terms glm for E(A | W)
               SL_Qr = "SL.npreg", # use kernel regression to guard against
                                   # misspecification of outcome regression
               SL_gr = "SL.npreg", # use kernel regression to guard against
                                  # misspecification of propensity score
               returnModels = TRUE # for visualizing fits later
              )
# print the output
fit1

# get confidence intervals for marginal means
# truth is E[Y(1)] = 0.29, E[Y(0)] = 0.15
ci_fit1 <- ci(fit1)
# print the output
ci_fit1

# get confidence intervals for ate
# truth is E[Y(1)] - E[Y(0)] = 0.14
ci_ate1 <- ci(fit1, contrast = c(-1, 1))
# print the output
ci_ate1
```

This method requires estimation of additional univariate regressions to ensure doubly robust confidence intervals and hypothesis tests. The method for estimation are input via `SL.Qr` and `SL.gr` or `glm.Qr` and `glm.gr` if parametric models are desired). These additional fits can be visualized by the `plot` method for `drtmle`. 
```r
layout(t(1:3))
plot(fit1, ask = FALSE)
```

### Inference for super learner-based IPTW

The package additionally includes a function for computing valid confidence
intervals about an inverse probability of treatment weight (IPTW) estimator when
super learning is used to estimate the propensity score.

```{r}
# fit iptw
fit2 <- adaptive_iptw(Y = Y, A = A, W = W, a_0 = c(0, 1),
                      SL_g = c("SL.glm", "SL.mean", "SL.step.interaction"),
                      SL_Qr = "SL.npreg")
# print the output
fit2

# compute a confidence interval for margin means
ci_fit2 <- ci(fit2)
# print the output
ci_fit2

# compute a confidence interval for the ate
ci_ate2 <- ci(fit2, contrast = c(-1, 1))
# print the output
ci_ate2
```
---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/benkeser/drtmle/issues).

---

## Citation

After using the `drtmle` R package, please cite the following:

    @Manual{drtmlepackage,
      title = {drtmle: Doubly-Robust Nonparametric Estimation and Inference},
      author = {David Benkeser},
      note = {R package version 1.0.0},
      doi = {10.5281/zenodo.844836},
    }

    @article{benkeser2017improved,
      year  = {2017},
      author = {Benkeser, David C and Carone, Marco and van der Laan, Mark J
        and Gilbert, Peter B},
      title = {Doubly-robust nonparametric inference on the average
        treatment effect},
      journal = {Biometrika},
      volume = {104}, number = {4},
      pages = {863–880},
      doi = {10.1093/biomet/asx053}
    }

---

## License

&copy; 2016-2017 [David C. Benkeser](http://www.benkeserstatistics.com)

The contents of this repository are distributed under the MIT license. See
below for details:
```
The MIT License (MIT)

Copyright (c) 2016-2017 David C. Benkeser

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```