### 2022-06-23 paper/using_drtmle_obstudies.Rmd
attempt paper conversion

---
documentclass: article
classoption: [twoside,11pt]
author: David Benkeser and Nima S. Hejazi
title: "Doubly-Robust Inference in R using \\texttt{drtmle}"
keywords: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
output:
  latex_document:
    extra_dependencies:
      obs_study_style: null
      amsmath: null
      mathtools: null
      fancyvrb: null
      inputenc: ["utf8"]
bibliography: refs.bib
header-includes: >
  \providecommand{\tightlist}{%
    \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
  \newcommand{\dataset}{{\cal D}}
  \newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
  \DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
  \DefineVerbatimEnvironment{Soutput}{Verbatim}{}
  \DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
  \newenvironment{Schunk}{}{}
  \DefineVerbatimEnvironment{Code}{Verbatim}{}
  \DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
  \DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
  \newenvironment{CodeChunk}{}{}
  \setkeys{Gin}{width=0.8\textwidth}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \texttt{drtmle} implements doubly-robust confidence intervals and hypothesis
  tests for targeted minimum loss-based estimates of the average treatment
  effect, in addition to several other recently-proposed estimators of the
  average treatment effect.
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
renv::activate(here::here())
library(nloptr); library(earth)
```

# Introduction

In many fields, researchers are interested in assessing the population-level
effect of a particular treatment on an outcome of interest. The treatment might
correspond to a drug, a potentially harmful exposure, or a policy intervention.
Often, the treatment cannot be randomized due to ethical or logistical reasons.
This has sparked great interest in developing statistical methodology to
estimate population-level effects from observational data. Estimation of these
effects often requires accounting for putative confounding factors, that is,
factors related to whether a participant receives treatment and to the
participant's outcome. In many settings, researchers measure many potential
confounders, for example using administrative databases or genetic sequencing
technology. Due to the curse of dimensionality, common statistical estimation
techniques are not feasible for such high-dimensional data without restrictive
modeling assumptions. When these assumptions are violated, estimates of the
population-level effect of a treatment may have large bias.

This has led to increased interest in adopting adaptive estimation techniques from the
machine learning literature to control for high-dimensional confounders when
estimating treatment effects. However, facilitating proper inference (i.e.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, the use of standard causal inference
techniques, including both G-computation (GCOMP) estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when utilizing
adaptive estimators. In particular, with rare exception, if cross-validation is used
to select tuning parameters for the nuisance parameters, estimates of the population-level
effect will generally be too biased to achieve standard $n^{1/2}$-rate asymptotics. Instead,
*undersmoothing* must be used to select tuning parameters to control the bias 
[@undersmoothed_hal]. However, this can prove quite challenging in practice and 
few guidelines are available.

A more natural framework for integrating adaptive methods into nuisance regression 
estimation are doubly robust approaches, including augmented inverse probability of 
treatment weighted (AIPTW) estimation [@robins:1994:jasa] and targeted minimum 
loss-based (TML) estimation [@vdlrubin:2006:ijb]. The asymptotics of these estimators
are characterized in part by the convergence rates of nuisance parameter estimates with
respect to a proper norm. This fact makes the use of cross-validation more appealing for
selection of tuning parameters for estimators of nuisance parameters, thereby providing 
a more straightforward means of obtaining an estimate of the effect of interest that
enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield closed form
standard error estimates, which can serve as the basis for constructing confidence 
intervals and hypothesis tests.

As an intermediate step, the AIPTW and TML estimators require estimation of
two key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and
TML estimators a natural appeal: inconsistency in the estimation of one
nuisance parameter may be mitigated by the consistent estimation of the other.
However, recent work has shown that the double robustness property does not
necessarily extend to the inferential properties of these estimators when
adaptive estimators of the PS and/or OR are used
[@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb
proposed new TML estimators that have an asymptotic normal sampling distribution
with estimable variance under consistent estimation of *either* the PS *or* the
OR, paving the way for inference that is doubly-robust.
@benkeser:2017:biometrika proposed modifications of these estimators with
similar robustness properties and illustrated their practical performance.

The two proposed procedures that yield doubly-robust inference are quite
involved, notably requiring iterative estimation of additional, complex nuisance
parameters. The \texttt{drtmle} package was motivated by the need for a
user-friendly tool to implement these methods. The package implements the TML
estimators of population-level effects proposed in @vdl:2014:ijb and
@benkeser:2017:biometrika, as well as several extensions including
cross-validated targeted minimum loss-based estimation (CV-TMLE). Additionally,
\texttt{drtmle} implements an IPTW estimator that overcomes shortcomings of
existing IPTW implementations with respect to statistical inference. In
particular, the estimator allows an adaptive estimator of the PS to be used in
construction of the IPTW estimator, while yielding valid statistical inference
about population-level effects. Both the TML and IPTW estimators implemented in
the \texttt{drtmle} package are capable of estimating population-level effects for
discrete-valued treatments; moreover, the package provides convenient utilities
for constructing confidence intervals and hypothesis tests about contrasts of
the mean outcome under different levels of treatment. This document explains
some of the key concepts underlying doubly-robust inference and illustrates
usage of the \texttt{drtmle} package.

# Definition and estimators of the average treatment effect

## Parameters of interest

Suppose the observed data consist of $n$ independent and identically
distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
a vector of baseline covariates, $A \in \{0,1\}$ is a binary treatment, 
$Y$ is an outcome, and $P_0$ is the true data-generating
distribution. The parameters of interest are
\begin{equation}
  \psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
  \mbox{for} \ a = 0,1 \ ,
  \label{parameterDef}
\end{equation}
where $\bar{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
assumptions, these parameters have interpretations as mean counterfactual
outcomes under the different treatments, with contrasts between these means
defining causal effects.

## G-computation estimators

An estimator of $\psi_0(a)$ may be motivated directly by (1). Suppose we have
available an estimate of the OR, $\bar{Q}_n$, and an estimate of the
distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
\bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
covariates is used so that this estimator can be equivalently expressed
\begin{equation*}
  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
\end{equation*}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

## Inverse probability of treatment estimators

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (1) can be equivalently written as
\begin{align}
  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
          \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation (2)
motivates the estimator
\begin{equation*}
  \psi_{n,IPTW}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap.
If adaptive approaches are used to estimate the PS, inference may not be
available without further modification of the propensity score estimator, as we
discuss below.

## Doubly-robust estimators

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness. That is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators generally lack a basis for statistical inference.

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
\psi_{n,AIPTW}(a) = \psi_{n,IPTW}(a) + \frac{1}{n} \sum\limits_{i=1}^n \biggl\{
1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \  \bar{Q}_n(a, W_i) \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
\psi_{n,AIPTW}(a) = \psi_{n,GCOMP}(a) + \frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. The estimator exhibits double-robustness in that it is consistent
for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
is consistent for the true PS. Furthermore, if both the OR and PS are
consistently estimated and regularity conditions hold, one can establish a
limiting normal distribution for the AIPTW estimator. Simple estimators of the
variance of this limiting distribution can be derived that may be used to
construct Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss-based estimation
(TMLE). TMLE is a general methodology that may be used to modify an estimator
of a regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\bar{Q}_n^{\star}$ satisfies
\begin{equation}\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \bar{Q}_n^{\star}(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly-robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies (4). In addition to possessing the desirable asymptotic
properties, TML estimators have been shown to outperform AIPTW estimators in
finite samples [@porter:2011:ijb]. We refer readers to @tlbook:2011 for
more about TMLE.

# Doubly-robust inference

@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. @vdl:2014:ijb derived
modified TML estimators that are doubly-robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly-robust variance estimators that may be used to
construct doubly-robust confidence intervals and hypothesis tests.
@benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
estimators and demonstrated the practical performance of these estimators via
simulation.

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (4) in addition to two additional
equations. These equations involve additional regression parameters that we
refer to as the reduced outcome regression (R-OR) and the reduced propensity
score (R-PS), defined respectively as
\begin{align}
  \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qbarr} \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
  g_n(W)=g_n(w)\right\} \ . \label{eqn:gr_bivariate}
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @{vdl:2014:ijb showed that if estimates of the
OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \bar{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \bar{Q}_n^{\star}(a, W_i) \} = 0 \ ,
\label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly-robust not
only with respect to consistency, but also with respect to its limiting
distribution. The author additionally proposed a TMLE algorithm to map initial
estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy these key
equations.

@benkeser:2017:biometrika demonstrated alternative equations that can be
satisfied to achieve this form of double-robustness. In particular, these
authors formulated an alternative version of the R-PS,
\begin{align*}
  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \bar{Q}_n(W) =
  \bar{Q}_n(w) \} \ ,
\end{align*}
which we refer to as R-PS1. They also introduced an additional regression of a
weighted residual from the PS on the OR,
\begin{align*}
  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
  {g_n(a \mid W)} \ \bigg\vert \ \bar{Q}_n(W) = \bar{Q}_n(w) \biggr\} \ ,
\end{align*}
which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
both univariate regressions, while R-PS is a bivariate regression.
@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
easier to estimate consistently than R-PS. Further, they showed that if
estimators of the OR, PS, and R-OR satisfy equations (5)-(6), and if,
additionally, estimators of the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2
$g_{r,n,2}^{\star}$ satisfy
\begin{equation*}
\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
\bar{Q}_n^{\star}(a,W_i)\} = 0 \ ,
\end{equation*}
then the GCOMP estimator based on this $\bar{Q}_n^{\star}$ also enjoys a
doubly-robust limiting distribution. @benkeser:2017:biometrika provided a
TMLE algorithm that produced such estimates and provided closed-form,
doubly-robust variance estimates.

\textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
methods for constructing doubly-robust (with respect to consistency) estimators
of our parameter of interest. Theoretically, these two frameworks are closely
related and, practically, the two estimators are generally seen as competitors
for estimating marginal means. In equation (3), we were able to generate a
doubly-robust estimator by adding a term to the GCOMP estimator. This term is
exactly the left-hand-side of the key equation (4), the key equation that is
solved by the standard TMLE. This is no coincidence; this term is crucial in
studying the asymptotic properties of the GCOMP, AIPTW, and TML estimators
(for discussion, see @benkeser:2017:biometrika Section 2.2). Recall that
a TML estimator achieves doubly-robust consistency by satisfying equation (4),
while the AIPTW achieves doubly-robust consistency by combining the GCOMP
estimator with the left-hand-side of this term. Now, we have argued that a TML
estimator may achieve doubly-robust limiting behavior if it additionally
satisfies equations (6) and (7). It is thus sensible to wonder whether an
AIPTW-style estimator could be constructed by combining the left-hand-side of
these two additional equations with the AIPTW estimator.
@benkeser:2017:biometrika showed that, surprisingly, this AIPTW-style estimator
does not achieve the same doubly-robust properties as the TML estimator. Thus,
while AIPTW and TMLE are generally competitors for producing estimators
doubly-robust with respect to consistency, for doubly-robust inference, TMLE
ought to be preferred.

# Implementation of doubly-robust inference

The main function of the package is the eponymous `drtmle()` function. This
function estimates the treatment-specific marginal mean for user-specified
levels of a discrete-valued treatment and computes a doubly-robust covariance
matrix for these estimates. The function implements either the
estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the
improved estimators of @benkeser:2017:biometrika
(`reduction = "univariate"`). The package includes utility functions for
combining these estimates into estimates of average treatment effects, as well
as other contrasts, as discussed below.

The main data arguments for `drtmle()` are `W, A, Y`, which, as above, represent
the covariates, a discrete-valued treatment, and an outcome, respectively.
Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
and R-PS (or R-PS1, R-PS2). The package provides flexibility in how these
regression parameters may be estimated. Once the user has specified the data
arguments and how to estimate each regression parameter, the function proceeds
as follows:
  1. estimate the OR via the internal `estimateQ()` function;
  2. estimate the PS via the internal `estimateG()` function;
  3. estimate reduced-dimension regression via the internal `estimateQrn()` and
     `estimategrn()` functions;
  4. iteratively modify initial estimates of the OR and PS via the
     `fluctuateQ()` and `fluctuateG()` functions until equations (5)-(7) are
     approximately solved;
  5. compute the TML estimates and covariance estimates based on the final OR.
We refer interested readers to @benkeser:2017:biometrika for a
theoretical derivation of how the iterative modification of initial OR and PS is
performed.

In addition to returning doubly-robust parameter and covariance estimates, the
`drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
modified AIPTW estimator discussed in the remark above. While doubly-robust
inference based on these estimators is not available, the estimators are
provided for comparison. In the following subsections, we look at various
options for making calls to `drtmle()`.

## Estimating regressions
```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
```
While the estimators computed by `drtmle()` are consistent and asymptotically
normal under inconsistent estimation of either the OR or PS, it is nevertheless
advisable to strive for consistent estimation of *both* the OR and PS. If both
of the regression parameters are consistently estimated, then the estimators
computed by `drtmle()` achieve the nonparametric efficiency bound, resulting in
narrower confidence intervals and more powerful hypothesis tests. In order for
the estimates of the OR and PS to have the greatest chance of consistency, the
`drtmle()` function facilitates estimation using adaptive estimation techniques.
In fact, there are three ways to estimate each of the regression parameters:
generalized linear models, super learning
[@vdl:2007:statapp; @SuperLearnerPackage], and a user-specified estimation
technique. We demonstrate each of these approaches in turn on a  simulated
data set. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0,
2)` and $\psi_0(1)=$ `r round(EY1, 2)`.

```{r}
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 1, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
```

### Using generalized linear models to estimate regressions

Generalized linear models are perhaps the most popular means of estimating
regression functions. Because of this legacy, these estimators have been
included as an option for estimating the regression parameters. The user
specifies the right-hand-side of a regression formula as a character in the
`glm_g`, `glm_Q`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS, R-OR,
and R-PS. These regressions are fit by calling the `glm()` function in R [@R].
Thus, the model specification should be able to be parsed as a valid formula,
as required by `glm()`. For the PS formula specified by `glm_g`, `colnames(W)`
are available to use in the formula. For the OR (`glm_Q`) available names
of variables to be used in the formula depends on the `stratify` option.
If `stratify = FALSE`, a single OR is fit to estimate $\bar{Q}_0(A,W)$. In this
case, the OR regression is fit by passing a `data.frame` that contains `W` and
`A` into the `data` argument of `glm()`. Thus, both `colnames(W)` and `"A"` may
appear in the formula specified in `glm_Q`. If `stratify = TRUE`, regressions
are fit stratifying on `A` to separately estimate $\bar{Q}_0(1,W)$ and
$\bar{Q}_0(0,W)$. In this case, the `glm_Q` formula may not include `"A"`.

To understand the variable names available for R-OR, recall from
(\ref{eqn:qbarr}) that the parameter involves a regression onto the estimated
propensity score. Thus, the formula that is input via `glm_Qr` should assume
data with a single column, `"gn"`.  Similarly, the R-PS1 and R-PS2 that are
computed for the estimators of @benkeser:2017:biometrika (the default
estimators, computed whenever `reduction = "univariate"`) involve regressions
onto the estimated outcome regression. Thus, the formula `glm_gr` should assume
data with a single column named `"Qn"` if `reduction = "univariate"`. Instead,
if the estimators of @vdl:2014:ijb are preferred (`reduction = "bivariate"`) the
`glm_gr` formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes data
with two columns, `"Qn"` and `"gn"`.

The code below illustrates how generalized linear models can be used to
estimate the regression parameters. In this call to `drtmle()`, we specify
`family = binomial()` so that logistic regression is used to estimate the OR.
We illustrate non-stratified regression for both the univariate (default) and
bivariate reductions below.

```{r}
glm_fit_uni <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE)
glm_fit_uni

glm_fit_biv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                      glm_gr = "Qn + gn", glm_Qr = "gn", stratify = FALSE,
                      reduction = "bivariate")
glm_fit_biv
```

### Using super learner to estimate regressions

```{r, echo = FALSE, message = FALSE}
library(SuperLearner)
```
Super learning is a loss-based ensemble machine learning technique
[@vdl:2007:statapp] that generalizes stacking and stacked regression
\citep{wolpert:1992:nnet,breiman:1996:ml}. The user specifies many potential
candidate regression estimators, referred to as a *library* of candidate
estimators, and the super learner algorithm estimates the convex combination of
regression estimators that minimizes the $V$-fold cross-validated risk based on
a user-selected loss function [@dudoit2005asymptotics].  Oracle inequalities
show that the super learner estimate performs essentially as well as the unknown
best convex combination of the candidate estimators [@vdvaart:2006:statdec].
Thus, the methodology may be seen as an asymptotically optimal way of performing
estimator selection. Practically, the super learner provides the best chance
for obtaining a consistent estimator for both the OR and PS, and thereby an
efficient estimator of the marginal means of interest.

Super learning is implemented in the R package \texttt{SuperLearner}
[@SuperLearnerPackage] and `drtmle()` provides the option to utilize super
learning for estimation of the OR, PS, and reduced-dimension regressions via the
`SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle()` either the
generalized linear model options for regression modeling or the super learner
options should be invoked. If both are called, the function will automatically
opt to using the super learner option. The inputs to the super learner options
are passed to the `SL.library` argument of the `SuperLearner()` function of the
\texttt{SuperLearner} package. We refer readers to the documentation of the
`SuperLearner()` function for information on how to properly specify a super
learner library. In the code below, we illustrate a call to `drtmle()` with a
relatively simple library. For the OR and PS, the library includes a main-terms
generalized linear model (via the function `SL.glm` from \texttt{SuperLearner}) and
an unadjusted generalized linear model (`SL.mean`). For the reduced-dimension
regressions we use a super learner library with a main-terms generalized linear
model (`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
\texttt{earth} package [@earthPackage]).

```{r, warning = FALSE, message = FALSE, cache = TRUE}
set.seed(1234)
sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                 SL_g = c("SL.glm","SL.mean"),
                 SL_Q = c("SL.glm","SL.mean"),
                 SL_gr = c("SL.glm", "SL.earth"),
                 SL_Qr = c("SL.glm", "SL.earth"),
                 stratify = FALSE)
sl_fit
```

The \texttt{drtmle} package also allows users to pass in their own functions for
estimating regression parameters via the `SL_` options. These functions must be
formatted properly for compatibility with `drtmle()`. The necessary formatting
is identical to the formatting required for writing a `SuperLearner` wrapper
(see `SuperLearner::write.SL.template()`). The \texttt{drtmle} package includes
`SL.npreg`, which is an example of how a user can specify a function for
estimating regression parameters. This functions fits a kernel regression
estimator using the \texttt{np} package [@npPackage]. Below we show the source code
for this function.

```{r}
SL.npreg
```

In the above code, we see that a user-specified function should have options
`Y, X, newX, family, obsWeights, ...`, along with other options specific to the
function. The option `Y` corresponds to the outcome of the regression and `X`
to the variables onto which the outcome is regressed. The option `newX` should
be of the same class and dimension as `X` and is meant to contain new values of
variables at which to evaluate the regression fit. The `family` and
`obsWeights` options are not strictly necessary (indeed, they are not used by
`SL.npreg`), though these may be useful to include if one wishes to utilize the
function as part of a super learner library. The `SL.npreg` function proceeds by
checking whether there is sufficient variation in `Y` to justify fitting the
kernel regression (as specified by option `rangeThresh`), and if so, fits a
kernel regression with cross-validated bandwidth selection via the `npregbw()`
function. The output is then formatted in a specific way so that `drtmle()` is
able to retrieve the appropriate objects from the fitted regression. In
particular the output should be a list with named objects `pred` and `fit`; the
former including predictions made from the fitted regression on `newX`, the
latter including any information that may be needed to predict new values based
on the fitted object. A class is assigned to the `fit` object, so that an S3
`predict()` method may later be used to predict new values based on the fitted
regression object. The user must also specify this `predict()` method. Use
`drtmle:::predict.SL.npreg()` to view the `predict()` method for
`SL.npreg`.

In the code below, we demonstrate a call to `drtmle()` with `SL.npreg` used to
estimate each nuisance parameter. Here, we also illustrate the use of `stratify
= TRUE`, which fits a separate kernel regression of $Y$ onto $W$ in
observations with $A = 1$ and $A = 0$.

```{r, cache = TRUE, message = FALSE}
npreg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = "SL.npreg", SL_Q = "SL.npreg",
                    SL_gr = "SL.npreg", SL_Qr = "SL.npreg",
                    stratify = TRUE)
```

In the Appendix, we describe an additional means of using `drtmle()` with 
OR and PS estimates that have been obtained externally.

### Reducing dependence on random seed

Some users may find it unsettling that when applying the super learner in real
data analysis the results are dependent on the random seed that is set. To
remove this dependence, one can average results over repeated super learner
runs; more repeats will lead to less dependence on the random seed. There are
generally two approaches that can be used: (i) averaging on the scale of the
nuisance parameters and (ii) averaging on the scale of the point estimates.

For (i), suppose that we have $n_{SL}$ estimated super learners for the OR,
$\bar{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS $g_{n,j}, j = 1, \dots,
n_{SL}$. We can take the average over these fits,
\begin{equation*}
  \bar{Q}_n = \frac{1}{n_{SL}} \sum_{j = 1}^{n_{SL}} \bar{Q}_{n,j}
\end{equation*}
as our estimate of the OR (similarly for the PS). These averaged estimates are
then used to compute the AIPTW or TMLE, as described above.

For (ii), suppose for example that we have $n_{SL}$ TML (similarly, AIPTW)
estimates $\psi_{n,j}(1)$ of $\psi_0(1)$, each obtained based on one (or more)
super learners of the OR and/or PS. We can take the average over these point
estimates,
\begin{equation*}
  \psi_n(1) = \frac{1}{n_{SL}} \sum_{j=1}^{n_{SL}} \psi_{n,j}(1)
\end{equation*}
as our estimate of $\psi_{0}(1)$.

This feature is implemented via the `n_SL` and `avg_over` options, where the
former specifies the number of super learners to repeat and the latter specifies
the scale(s) to average over. For example, if `avg_over = "drtmle"` and `n_SL =
2`, then two point estimates are computed, each based on a single super learner,
and averaged to obtain a final estimate. If instead `avg_over = "SL"` and `n_SL
= 2`, then two super learners are fit and averaged before a final point estimate
is obtained. Finally, if `avg_over = c("drtmle", "SL")` and `n_SL = 2` then
averaging on *both* scales is performed. That is, the final estimate is the
average of two point estimates, each obtained using OR and PS estimates that are
averaged over two super learners.

Not surprisingly, averaging over multiple super learners can add considerable
computational expense to calls to `drtmle()`. The package is currently
parallelized at the level of individual calls to `SuperLearner()`. With
parallelization over enough cores, repeated super learning should not add too
much to the computational burden relative to a single call to `SuperLearner()`.

## Standard error estimation

This section contains some more advanced mathematical material and may be
skipped over by users interested only in implementation. For simplicity, we
focus this discussion on the standard TML and AIPTW estimators. The ideas
generalize immediately to the DR-TML estimator, but the formulas are more
complex in this latter case.

Reported standard error estimates are based on an estimate of the asymptotic
variance of the estimators. This variance can be characterized by influence
function of the estimators. For example, let $\psi_{n}(1)$ be an AIPTW estimate
of $\psi_0(1)$. Under regularity conditions,
\begin{equation*}
  \psi_{n}(1) - \psi_0(1) = \frac{1}{n} \sum_{i=1}^n
    D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
\end{equation*}
where
\begin{equation*}
  D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}
  {g_0(1 \mid W_i)} \{ Y_i - \bar{Q}_0(1, W_i)\} + \bar{Q}_0(1, W_i) -
  \int \bar{Q}_0(1, w) dQ_{W,0}(w) \ .
\end{equation*}
Thus, by the central limit theorem, $n^{1/2} \psi_{n}(1)$ converges in
distribution to a normal distribution with mean $\psi_0(1)$ and variance
\begin{equation*}
  \sigma^2_0 = E_0( [ D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O) -
  E_0\{D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
\end{equation*}
An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical
variance of $D_i = D^{\star}(\bar{Q}_n, g_n, Q_{W,n})(O_i)$, that is,
\begin{equation*}
  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{D_i - \frac{1}{n} \sum_{j=1}^n
  D_j \right\}^2 \ .
\end{equation*}
Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n /
n^{1/2}$. Conveniently, this idea generalizes naturally to estimating the
asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.

### Standard errors for TMLE

For TML estimators, it is standard practice to construct $D_i$ using the
*targeted* version of the OR (denoted by $\bar{Q}_n^{\star}$ above). Recall that
such estimates are obtained by modifying an initial OR estimate so as to ensure
a particular (set of) equation(s) is (are) solved.  However, simulations have
shown that better standard error estimates for TMLE may be achieved by instead
using the *initial* OR estimate in the computation of standard errors, as
described in the preceding subsection. To this end, The `drtmle()` function
includes the `targeted_se` option.

```{r, show_targeted_se, cache = TRUE}
glm_fit_uni_nontargeted_se <-
  drtmle(W = W, A = A, Y = Y, family = binomial(),
         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
         glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE,
         targeted_se = FALSE)
# compare to original call
list(
  glm_fit_uni$drtmle$cov, # based on targeted OR
  glm_fit_uni_nontargeted_se$drtmle$cov # untargeted OR
)
```

### Cross-validated standard errors

Simulation studies have revealed that, when employing machine learning to
estimate the OR and/or PS, the estimated standard errors are often *too small*,
in small to moderate-sized samples. Problematically, this leads directly to
under-coverage of confidence intervals and possibly inflated Type-I error rates
of hypothesis tests based on these standard error estimates. This should not be
too surprising, as the phenomenon is related to the form of the standard error
estimates. In particular, such estimates include a term that resembles a
(weighted) mean squared error (MSE) of the OR. Since machine learning algorithms
often set out to minimize empirical MSE as part of their training process, it
follows that the empirical MSE often ends up biased towards zero relative to the
true MSE. To more accurately estimate the MSE of a machine learning algorithm,
cross-validation can be used. The same sample-splitting idea can be used to
obtain more conservative standard error estimates.

Suppose we use $V$-fold cross-validation to obtain $V$ training-sample-specific
estimates of the OR, $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v},
v = 1, \dots, V$. For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the
estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the
validation fold. That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of
$Y_i$. We define $g_{n,i}^{cv}$ similarly. Using these estimates, we can define
\begin{equation*}
 D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} +
 \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1)
\end{equation*}
and compute an estimate of $\sigma^2_0$ using
\begin{equation*}
  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} -
  \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ .
\end{equation*}

In \texttt{drtmle}, cross-validated standard error estimates can be obtained by
setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1. For
example,

```{r, fit_w_full_cv, warning = FALSE, cache = TRUE}
glm_fit_uni_cv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         se_cv = "full", se_cvFolds = 2, stratify = FALSE,
                         targeted_se = FALSE) # needed for this to work
# compare variance to previously obtained
list(
  `No CV` = glm_fit_uni$drtmle$cov, 
  `CV` = glm_fit_uni_cv$drtmle$cov
)
```

### Partially cross-validated standard errors

Obtaining fully cross-validated standard errors can be fairly time-intensive
when super learner is used to estimate the OR and PS. In \texttt{drtmle}, we give
an option to compute standard error estimates that are *partially
cross-validated* and do not require any additional fitting beyond a standard
super learner. Here, we describe briefly how these estimates are obtained and
show calls to `drtmle()` for obtaining such estimates.

Suppose we have $K$ candidate regressions in our super learner library for both
the OR and PS (in practice, the number could be different for the OR and for the
PS). Recall that a super learner estimate of, for example, $\bar{Q}_0(a, w)$ is
obtained as a weighted combination of estimates from each candidate regression.
Let $\bar{Q}_{n,k}$ denote the estimate of $\bar{Q}_{0}$ obtained by training
algorithm $k$ using the full data set. The super learner estimate of
$\bar{Q}_{0}(a,w)$ is
\begin{equation*}
  \bar{Q}_{n,SL}(a, w) = \sum_{k = 1}^K \alpha_{n,k} \bar{Q}_{n,k}(a, w) \ ,
\end{equation*}
where $\alpha_n = (\alpha_{n,1}, \dots, \alpha_{n,K})$ is a weight vector that
is selected by minimizing a cross-validated risk criterion. That is, each of the
$K$ algorithms is trained according to a $V$-fold cross-validation scheme,
resulting in $V + 1$ fits of each algorithm (one in each of $V$ training folds,
plus one obtained by fitting the algorithm using the full data). Let
$\bar{Q}_{n,k,v}$ denote the $k$\textsuperscript{th} algorithm trained using the
$v$\textsuperscript{th} training fold. For $i = 1, \dots, n$, let
$\bar{Q}_{n,k,i}^{cv}(1)$ denote the estimate of $\bar{Q}_0(1, W_i)$ implied by
the fitted algorithm $\bar{Q}_{n,k,v}$, which was obtained when observation $i$
was in the validation sample. That is, for observations with $A_i = 1$,
$\bar{Q}_{n,k,i}^{cv}(1)$ is an out-of-sample prediction of $Y_i$ based on the
$k$\textsuperscript{th} algorithm. Now, we can construct a *partially
cross-validated* super learner prediction as follows. For $i = 1, \dots, n$ let
\begin{equation*}
  \bar{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k}
    \bar{Q}_{n,k,i}^{cv}(1)
\end{equation*}
be the estimate obtained by ensembling the training-set-specific algorithm fits
based on $\alpha_n$. We can use these estimates to produce our *partially
cross-validated* standard error estimate as follows. First, define
\begin{equation*}
  D_i^{pcv} = \frac{I(A_i = 1)}{g_{n,SL,i}^{cv}} \{ Y_i -
    \bar{Q}_{n,SL,i}^{cv}(1)\} + \bar{Q}_{n,SL,i}^{cv}(1) -
    \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,SL,i}^{cv}(1),
\end{equation*}
then, compute an estimate of $\sigma^2_0$ using
\begin{equation*}
  \sigma^2_{n,pcv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{pcv} -
  \frac{1}{n} \sum_{j=1}^n D_j^{pcv} \right\}^2 \ .
\end{equation*}

Note that is not a properly cross-validated estimate, since observation $i$ is
used to compute the weight vector $\alpha_n$. However, such estimates can be
obtained based on the output of a single round of super learning, thereby
providing improved standard error estimates at minimal additional computational
cost.

Below, we demonstrate how to obtain such estimates using `drtmle()`.

```{r, demo_pcv, cache = TRUE, warning = FALSE}
set.seed(1234)
sl_fit_pcv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm","SL.mean"),
                     SL_Q = c("SL.glm","SL.mean"),
                     SL_gr = c("SL.glm", "SL.earth"),
                     SL_Qr = c("SL.glm", "SL.earth"),
                     stratify = FALSE,
                     se_cv = "partial")
# compare estimates
# pt estimates should be same
# variance should be larger for pcv
list(
  `No-CV` = sl_fit$cov, # no cv
  `Partial CV` = sl_fit_pcv$cov # partial cv
)
```

## Confidence intervals

The \texttt{drtmle} package contains methods that compute doubly-robust confidence
intervals. The `ci()` method computes confidence intervals about the estimated
marginal means in a fitted object of S3 class `"drtmle"`, in addition to
contrasts between those means. By default, the method gives confidence intervals
about each mean.

```{r}
ci(npreg_fit)
```

Alternatively the `contrast` option allows users to specify a linear
combination of marginal means, which can be used to estimate the average
treatment effect, as we show below.

```{r}
ci(npreg_fit, contrast = c(1, -1))
```

The `contrast` option also accepts as input a list of functions to be used to
compute a confidence interval for a user-specified contrast. For example, we
might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
constructing Wald-style confidence intervals for ratios, it is common to
compute the interval on the logarithmic scale and subsequently back-transform.
That is, the confidence interval takes the form
\begin{equation*}
  \exp \biggl[\log \biggl\{ \frac{\psi_n(1)}{\psi_n(0)} \biggr\} \pm
  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
\end{equation*}
where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
the delta method, an estimate of this standard error is given by
\begin{equation*}
  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
\end{equation*}
where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
`drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
This confidence interval may be obtained as follows.

```{r}
riskRatio <- list(f = function(eff){ log(eff) },
                  f_inv = function(eff){ exp(eff) },
                  h = function(est){ est[1] / est[2] },
                  fh_grad =  function(est){ c(1 / est[1], -1 / est[2]) })
ci(npreg_fit, contrast = riskRatio)
```

More generally, this manner of inputting the `contrast` option to the `ci()`
method can be used to compute confidence intervals of the form
\begin{equation*}
  f^{-1}\bigl\{ f(h(\psi_n)) \pm z_{1-\alpha/2} \nabla f(h(\psi_n))^T \Sigma_n
  \nabla f(h(\psi_n)) \bigr\} \ ,
\end{equation*}
where $f$ (`contrast$f()`) is the transformation of the confidence interval,
$f^{-1}$ (`contrast$f_inv()`) is the back-transformation of the interval, $h$
(`contrast$h()`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
(`contrast$fh_grad()`) defines the gradient of the transformed contrast at the
estimated marginal means.

We note that this method of computing confidence intervals can also be used to
obtain a transformed confidence interval for a particular marginal mean. For
example, recall that $Y$ is binary in our running example. Thus, we may wish to
enforce that our confidence interval be computed on a scale ensuring that the
estimated confidence limits fall within the unit interval $[0, 1]$. To that end,
we can utilize an interval of the form:
\begin{equation*}
  \text{expit} \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
  \pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
\end{equation*}
This confidence interval may be implemented as follows.

```{r}
logitMean <- list(f = function(eff){ qlogis(eff) },
                  f_inv = function(eff){ plogis(eff) },
                  h = function(est){ est[1] },
                  fh_grad = function(est){ c(1 / (est[1] - est[1]^2), 0) })
ci(npreg_fit, contrast = logitMean)
```

## Hypothesis tests

Doubly-robust Wald-style hypothesis tests may be implemented using the
`wald_test()` method. As with confidence intervals, there are three options for
testing. First, we may test the null hypothesis that the marginal means equal a
fixed value (or values). Below, we test the null hypothesis that $\psi_0(1) =
0.5$ and that $\psi_0(0) = 0.6$ by inputting a vector to the `null` option.

```{r}
wald_test(npreg_fit, null = c(0.5, 0.6))
```

As with the `ci()` method, the `wald_test()` method allows users to test linear
combinations of marginal means by inputting a vector of ones and negative ones
in the `contrast` option. We can use this to test the null hypothesis of no
average treatment effect or to test the null hypothesis that the average
treatment effect equals a given reference value, e.g., $0.05$, by specifying
the `null` option.

```{r}
wald_test(npreg_fit, contrast = c(1, -1))
wald_test(npreg_fit, contrast = c(1, -1), null = 0.05)
```

The `wald_test()` method also allows for user-specified contrasts to be tested
using syntax similar to the `ci()` method. The only difference syntactically is
that it is not strictly necessary to specify `f_inv()`, as the hypothesis test is
performed on the transformed scale. That is, we can generally test the null
hypothesis that $f(h(\psi_0))$ equals $f_0$ (the function $f$ applied to the
value passed to `null`) using the Wald statistic
\begin{equation*}
  Z_n \coloneqq \frac{\{f(h(\psi_n)) - f_0\}}{\nabla f(h(\psi_n))^T \Sigma_n
  \nabla f(h(\psi_n))} \ .
\end{equation*}
We can use the `riskRatio` contrast defined previously to test the null
hypothesis that $\psi_0(1) / \psi_0(0) = 1$.

```{r}
wald_test(npreg_fit, contrast = riskRatio, null = 1)
```

## Missing data

The `drtmle()` function supports missing data in `A` and `Y`. Such missing data
are common in practice and, if ignored, can bias estimates of the marginal
means of interest. The estimators previously discussed can be modified to allow
such missingness processes to depend on observed covariates. Let $\Delta_A$ and
$\Delta_Y$ be indicators that $A$ and $Y$ are observed, respectively. We can
redefine the OR to be $\bar{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
A = a, \Delta_A = 1, \Delta_Y = 1, W = w)$ and similarly redefine the PS to be
\begin{align}
  g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
  \notag \\
  &= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
  \label{modPS} \\
  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w) \
   \notag.
\end{align}

To demonstrate, we artificially introduce missing values to `A` and `Y` in our
running example.

```{r}
DeltaA <- rbinom(n, 1, plogis(2 + W$W1))
DeltaY <- rbinom(n, 1, plogis(2 + W$W2 + A))
A[DeltaA == 0] <- NA
Y[DeltaY == 0] <- NA
```

The only modification of the call to `drtmle()` is in how the PS estimation is
performed. The options `glm_g` and `SL_g` are still used to fit generalized
linear models or a super learner, respectively. However, the code allows for
separate regression fits for each of the three components of the PS in (8). To
perform separate generalized linear model fits for each of the three components,
`glm_g` should be a named list with entries `"DeltaA"`, `"A"`, and `"DeltaY"`.
Respectively, these should provide a formula expression for the regression of
$\Delta_A$ on $W$, $A$ on $W$ in observations with $\Delta_A = 1$, and of
$\Delta_Y$ on $W$ and $A$ in observations with $\Delta_A = \Delta_Y = 1$ (if
`stratify = FALSE`) _or_ of $\Delta_Y$ on $W$ to be fit in observations with
$\Delta_A = \Delta_Y = 1$ and $A = a$ for each $a$ specified in option `a_0` (if
`stratify = TRUE`). If only a single formula is passed to `glm_g`, it is used
for each of the three regressions. We illustrate a call to `drtmle()` with
missing data and generalized linear model fits below.

```{r, cache = TRUE}
glm_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                       glm_g = list(DeltaA = "W1 + W2", A = "W1 + W2",
                                    DeltaY = "W1 + W2 + A"),
                       glm_Q = "W1 + W2*A", glm_Qr = "gn",
                       glm_gr = "Qn", family = binomial())
glm_fit_wNAs
```

It is possible to mix generalized linear models and super learning when fitting
each of the components of the PS. In the example below, we use the wrapper
function `SL.glm`, which fits a main terms logistic regression, to fit the
regression of $\Delta_A$ on $W$; use `SL.npreg` to fit the regression of $A$ on
$W$; and use a super learner with library `SL.glm`, `SL.mean`, and `SL.earth` to
fit the regression of $\Delta_Y$ on $W$ and $A$.

```{r, cache = TRUE}
mixed_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         SL_g = list(DeltaA = "SL.glm", A = "SL.npreg",
                         DeltaY = c("SL.glm", "SL.mean", "SL.earth")),
                         glm_Q = "W1 + W2*A", glm_Qr = "gn",
                         glm_gr = "Qn", family = binomial())
```

We can also fit the PS regressions externally and pass in via `gn`, though this
process is slightly more complicated than when no missing data are present. The
basic idea is to fit one regression for (i) the indicator of missing $A$, (ii)
$A$ itself, and (iii) the indicator of the outcome being missing. The three PS's
are then multiplied together and an appropriately ordered list is constructed as
before. We illustrate with the following example.

```{r, cache = TRUE}
# first regress indicator of missing A on W
fit_DeltaA <- glm(DeltaA ~ . , data = W, family = binomial())

# get estimated propensity for observing A
ps_DeltaA <- predict(fit_DeltaA, type = "response")

# now regress A on W | DeltaA = 1
fit_A <- glm(A[DeltaA == 1] ~ . , data = W[DeltaA == 1, , drop = FALSE],
             family = binomial())

# get estimated propensity for receiving A = 1
ps_A1 <- predict(fit_A, newdata = W, type = "response")
# propensity for receiving A = 0
ps_A0 <- 1 - ps_A1

# now regress DeltaY on A + W | DeltaA = 1. Here we are pooling over
# values of A (i.e., this is what drtmle does if stratify = FALSE).
# We could also fit two regressions, one of DeltaY ~ W | DeltaA = 1, A = 1
# and one of DeltaY ~ W | DeltaA = 1, A = 0 (this is what drtmle does if
# stratify = TRUE).
fit_DeltaY <- glm(DeltaY[DeltaA == 1] ~ . ,
                  data = data.frame(A = A, W)[DeltaA == 1, ],
                  family = binomial())

# get estimated propensity for observing outcome if A = 1
ps_DeltaY_A1 <- predict(fit_DeltaY, newdata = data.frame(A = 1, W),
                        type = "response")

# get estimated propensity for observing outcome if A = 0
ps_DeltaY_A0 <- predict(fit_DeltaY, newdata = data.frame(A = 0, W),
                        type = "response")

# now combine all results into a single propensity score
gn <- list(
  # propensity for DeltaA = 1, A = 1, DeltaY = 1
  ps_DeltaA * ps_A1 * ps_DeltaY_A1,
  # propensity for DeltaA = 1, A = 0, DeltaY = 1
  ps_DeltaA * ps_A0 * ps_DeltaY_A0
)

# pass in this gn to drtmle
out_fit_ps <-  drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                      # make sure a_0/gn are ordered properly!
                      gn = gn, a_0 = c(1, 0),
                      glm_Q = "W1 + W2*A", glm_Qr = "gn",
                      glm_gr = "Qn", family = binomial())
out_fit_ps
```

## Multi-level treatments

```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y2 <- rbinom(N, 1, plogis(W$W1 + 2*W$W2))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
EY2 <- mean(Y2)
```

In our examples so far, we have only considered binary treatments. However,
\texttt{drtmle} is fully equipped to handle treatments with an arbitrary number of
discrete values as well. Suppose that $A$ assumes values in a set $\mathcal{A}$
of limited cardinality, the PS regression estimation is modified to ensure that
\begin{equation*}
  \sum\limits_{a \in \mathcal{A}} g_n(a \mid w) = 1 \ \mbox{for all $w$} \ .
\end{equation*}
If this condition is true, we say that the estimates of the PS are compatible.
Many regression methodologies that have been developed work well with binary
outcomes, but have not been extended to multi-level outcomes. To ensure that
users of \texttt{drtmle} have access to the large number of binary regression
techniques when estimating the PS, we have taken an alternative approach to
estimation of the propensity for each level of the treatment. Specifically,
rather than fitting a single multinomial-type regression, we fit a series of
binomial regressions. As a concrete example, consider the case of no missing
data and where $\mathcal{A} = \{0,1,2\}$. We can ensure compatible estimates of
the PS by generating an estimate $g_n(0 \mid w)$ of $Pr_0(A = 0 \mid W = w)$
and $\tilde{g}_n(1 \mid w)$ of $Pr_0(A = 1 \mid A > 0, W = w)$. The latter
estimate may be generated by regressing $I(A = 1)$ on $W$ in observations with
$A > 0$. Since the outcome is binary, this regression may be estimated using
any technique suitable for binary outcomes. By rules of conditional
probability, it holds that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
Pr_0(A = 0 \mid W)\}$ and thus an estimate of $g_0(1 \mid w)$ can be computed
as $g_n(1 \mid w) = \tilde{g}_n(1 \mid w) \{1 - g_n(0 \mid w)\}$. Finally, we
let $g_n(2 \mid w) = 1 - g_n(0 \mid w) - g_n(1 \mid w)$, which ensures
compatibility of the estimates for each level of the covariates. This approach
generalizes to an arbitrary number of treatment levels. Below, we provide an
illustration using simulated data with a treatment that has three levels. The
true parameter values in this simulation are $\psi_0(0)=$ `r round(EY0,2)`,
$\psi_0(1)=$ `r round(EY1, 2)`, and $\psi_0(2)=$ `r round(EY2,2)`.

```{r, cache = TRUE, message = FALSE}
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 2, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
```

The multi-level PS can still be estimated using any of the three techniques
discussed previously. Here, we illustrate with generalized linear
models.

```{r, cache = TRUE, message = FALSE}
glm_fit_multiA <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         family = binomial(), a_0 = c(0,1,2))
glm_fit_multiA
```

Above, we used the option `a_0` to specify the levels of the treatment at which
we were interested in estimating marginal means. There is no need for `a_0` to
contain all unique values of `A`. For example, certain levels of the treatment
may not be of scientific interest, but nevertheless we would like to use these
data to help estimate the OR and PS (by setting `stratify = FALSE`).

The confidence interval and testing procedures extend to multi-level treatments
with minor modifications. We can still obtain a confidence interval and
hypothesis test for each of the marginal means by not specifying a `contrast`.

```{r, cache = TRUE}
ci(glm_fit_multiA)
wald_test(glm_fit_multiA, null = c(0.4, 0.5, 0.6))
```

Alternatively, we can specify a `contrast` to estimate confidence intervals
about the average treatment effect of $A=1$ versus $A=0$. Calls to `wald_test()`
can also be made.

```{r, cache = TRUE}
ci(glm_fit_multiA, contrast = c(-1, 1, 0))
```

Similarly, we can compute confidence intervals comparing the average treatment
effect of $A = 2$ versus $A = 0$.

```{r, cache = TRUE}
ci(glm_fit_multiA, contrast = c(-1, 0, 1))
```

User-specified contrasts are also available. For example, we can modify the
`riskRatio` object we used previously for a two-level treatment to compute the
risk ratio comparing $A=1$ to $A=0$ and comparing $A = 2$ to $A = 0$.

```{r, cache = TRUE}
riskRatio_1v0 <- list(f = function(eff){ log(eff) },
                      f_inv = function(eff){ exp(eff) },
                      h = function(est){ est[2]/est[1] },
                      fh_grad =  function(est){ c(1/est[2], -1/est[1], 0) })
ci(glm_fit_multiA, contrast = riskRatio_1v0)
```

Similarly, we can compute confidence intervals for the risk
ratio comparing the marginal mean for $A=2$ to $A=1$.

```{r, cache = TRUE}
riskRatio_2v0 <- list(f = function(eff){ log(eff) },
                      f_inv = function(eff){ exp(eff) },
                      h = function(est){ est[3]/est[1] },
                      fh_grad =  function(est){ c(0, -1/est[1], 1/est[3]) })
ci(glm_fit_multiA, contrast = riskRatio_2v0)
```

## Cross-validated regression estimates

To avoid overfitting initial regressions, one can use cross-validated 
estimates of the nuisance
regressions. Theoretically, the use of cross-validated regression weakens the
assumptions necessary to achieve asymptotic normality [@zheng:2010:working].
This is true of the regular TML and AIPTW estimators, as well as of the TML
estimator with doubly-robust inference. The `drtmle()` function implements
cross-validated estimation of the regression parameters through the `cvFolds`
option, as shown in the code below.

```{r, cache = TRUE}
cv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_Q = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_gr = c("SL.glm", "SL.mean"),
                    SL_Qr = c("SL.glm", "SL.mean"),
                    stratify = FALSE, cvFolds = 2, a_0 = c(0, 1, 2))
```

Cross-validation may significantly increase the computation time necessary for
running `drtmle()`. Parallelized nuisance regression fitting is available by
setting the option `use_future = TRUE`. Since, internally, all computations are
carried out using futures (with the API provided by the \texttt{future} R package),
setting this option amounts only to parallelizing with futures [@futurePackage]
via `future.apply::lapply()`; when `use_future = FALSE`, base `lapply()` is
simply used instead.  Together, setting these options results in parallelization
of the estimation of the OR, PS, and reduced-dimension regressions.

If no adjustments are made prior to invoking the `drtmle()` function, futures
are computed sequentially, as per the defaults of the \texttt{future} package. It
is left to the user to specify appropriate system-specific future backends that
may better suit their problem (e.g., using the \texttt{future.batchtools} package
[@futurebatchtoolsPackage] or even simply setting a parallelized
`future::plan()`). Below, we demonstrate registration using an ad-hoc cluster on
the local system. The choice of backend is flexible, and submission of the job
to a wide variety of schedulers (e.g., SLURM, TORQUE) is also permitted.
Interested readers are invited to consult the documentation of the
\texttt{future.batchtools} package for further information.

```{r, message = FALSE, eval = FALSE}
library(future.batchtools)
library(parallel)
cl <- makeCluster(2, type = "SOCK")
plan(cluster, workers = cl)
clusterEvalQ(cl, library("SuperLearner"))
pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
                     SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
                     SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
                     SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
                     stratify = FALSE, a_0 = c(0,1,2),
                     cvFolds = 2, use_future = TRUE)
```

# Discussion

The \texttt{drtmle} package was designed to provide a user-friendly implementation
of the TMLE algorithm that facilitates doubly-robust inference about marginal
means and average treatment effects. While the theory underlying doubly-robust
inference is complex, the implementation of the methods only requires users to
provide the data and specify how to estimate nuisance regressions. Though
estimation of the OR and PS is familiar to those with backgrounds in causal
inference-related estimation, the reduced-dimension regressions may appear
strange. Indeed, it is difficult to provide intuition for these parameters and
more difficult still to determine what scientific background knowledge might
guide users in how to estimate these regression parameters. Therefore, we
recommended estimating these quantities adaptively, e.g., with the super
learner. The super learner library should contain several relatively smooth
estimators such as `SL.mean`, `SL.glm`, and `SL.gam`. In particular, including
`SL.mean` (an unadjusted generalized linear model) may be important due to the
fact that, when the OR and PS are consistently estimated, the reduced-dimension
regressions are identically equal to zero. Thus, this unadjusted regression
estimator may do a good job estimating the reduced-dimension regressions in
situations where the OR and PS are estimated well. In addition to relatively
smooth estimators, we recommend additionally including several adaptive
estimators such as `SL.npreg` or `SL.earth`.

Planned extensions of the \texttt{drtmle} package include extensions to
longitudinal settings involving treatments at multiple time points and inclusion
of the collaborative targeted maximum likelihood estimator (C-TMLE) of
@vdl:2014:ijb that also facilitates collaborative doubly-robust inference.

# Appendix

## Arbitrary user-specified regressions

The \texttt{drtmle} package also allows users to pass in their own estimated OR and
PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
user, `drtmle()` will ignore the nuisance parameter estimation routines
specified by `SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As
noted in the package documentation, there is a specific format necessary for
inputting `Qn` and `gn`. For `Qn`, the entries in the list should correspond to
the OR evaluated at `A` and the observed values of `W`, with order determined by
the input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the
OR at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
example below illustrates this concept.

```{r, cache = TRUE, message = FALSE}
# use simplified data
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 1, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))

# fit a GLM for outcome regression outside of drtmle
out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())

# generate Qn list
Qn10 <- list(
  # first entry is predicted values setting A = 1
  predict(out_glm_fit, newdata = data.frame(A = 1, W), type = "response"),
  # second entry is predicted values setting A = 0
  predict(out_glm_fit, newdata = data.frame(A = 0, W), type = "response")
)

# pass this list to drtmle to avoid internal estimation of
# outcome regression (note propensity score and reduced-dimension
# regressions are still estimated internally)
out_fit1 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn", Qn = Qn10,
                  # crucial to set a_0 to match Qn's construction!
                  a_0 = c(1,0))
out_fit1

# to be completely thorough let's re-order Qn10 and re-run
Qn01 <- list(Qn10[[2]], Qn10[[1]])

out_fit2 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn",
                  # use re-ordered Qn list
                  Qn = Qn01,
                  # a_0 has to be reordered to match Qn01!
                  a_0 = c(0,1))
# should be the same as out_fit1, but re-ordered
out_fit2
```

Similarly, for `gn` we need to pass in properly ordered lists. For example, if
`a_0 = c(0, 1)` then `gn[[1]]` should be the PS for receiving $A = 0$ and
`gn[[2]]` should be the PS for receiving $A = 1$. The example below illustrates
this concept.

```{r, cache = TRUE, message = FALSE}
# fit a GLM for propensity score outside of drtmle
out_glm_fit <- glm(A ~ . , data = data.frame(W), family = binomial())

# get P(A = 1 | W) by calling predict
gn1W <- predict(out_glm_fit, newdata = data.frame(W), type = "response")

# generate gn list
gn10 <- list(
  # first entry is P(A = 1 | W)
  gn1W,
  # second entry is P(A = 0 | W) = 1 - P(A = 1 | W)
  1 - gn1W
)

# pass this list to drtmle to avoid internal estimation of
# propensity score (note reduced-dimension regressions are
# still estimated internally)
out_fit3 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                   glm_Qr = "gn", glm_gr = "Qn",
                   Qn = Qn10, gn = gn10,
                   # crucial to set a_0 to match Qn and gn's construction!
                   a_0 = c(1,0))
out_fit3

# to be completely thorough let's re-order gn10 and re-run
gn01 <- list(gn10[[2]], gn10[[1]])

out_fit4 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                   glm_Qr = "gn", glm_gr = "Qn",
                   # use re-ordered Qn/gn list
                   Qn = Qn01, gn = gn01,
                   # a_0 has to be reordered to match Qn01!
                   a_0 = c(0,1))
# should be the same as out_fit3, but re-ordered
out_fit4
```

This flexibility allows users to leverage any regression technique to externally
obtain initial estimates of nuisance parameters. The key step is making sure
that the ordering of `Qn` and `gn` is consistent with `a_0`.

## Additional methods included in drtmle

### Inference for IPTW using adaptive PS estimators

Doubly-robust estimators of marginal means are appealing in their robustness and
efficiency properties. However, researchers may prefer the IPTW estimator for
its intuitive derivation and implementation. Heretofore, inference for IPTW
estimators precluded the use of adaptive estimators of the PS. However,
@vdl:2014:ijb proposed modified IPTW estimators that allow for adaptive
PS estimation. Similarly as above, this estimator is based on an additional
regression parameter, $\tilde{Q}_{r,0n}(a,w) \coloneqq E_0\{Y \mid A = a, g_n(W)
= g_n(w)\}$, that is, the regression of the outcome $Y$ on the estimated
propensity score amongst observations with $A=a$. The author showed that if
$g_n^{\star}(a \mid w)$, an estimator of the propensity for treatment $A=a$,
satisfied that
\begin{equation*}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
    W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,
\end{equation*}
then the IPTW estimator based on $g_n^{\star}$ would have an asymptotic normal
distribution with estimable variance. In fact, @vdl:2014:ijb proposed a
TMLE algorithm that mapped an initial PS estimate into an estimate that
satisfies this key equation and variance estimators that can be used to generate
Wald-style confidence intervals and hypothesis tests.

An estimator with equivalent asymptotic properties based on a PS estimate,
$g_n(a \mid w)$, can be computed as
\begin{equation*}
  \psi_{n,IPTW}(a) = \psi_{n,IPTW}(a) - \frac{1}{n} \sum\limits_{i=1}^n
  \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)} \ \{ I(A_i = a) - g_n(a
  \mid W = W_i)\} \ .
\end{equation*}
Both this estimator and the TMLE-based IPTW estimator above are implemented in
\texttt{drtmle}.

#### Implementation of inference for adaptive IPTW

The IPTW estimators based on adaptive estimation of the PS are implemented in
the `adaptive_iptw()` function. The `adaptive_iptw()` function shares many
options with `drtmle()`; however, as the IPTW estimator does not rely on an
estimate of the OR, options related to the OR are omitted.

```{r, cache = TRUE}
npreg_iptw_multiA <- adaptive_iptw(W = W, A = A, Y = Y, stratify = FALSE,
                                   SL_g = "SL.npreg", SL_Qr = "SL.npreg",
                                   family = binomial(), a_0 = c(0, 1))
npreg_iptw_multiA
```

The `"adaptive_iptw"` S3 class contains identical methods for computing
confidence intervals and Wald-style tests as the `"drtmle"` class, thus readers
are referred back to the section on confidence intervals and hypothesis tests
for reminders on the various options for these procedures. By default, these
methods are implemented for the `iptw_tmle()` estimator of @vdl:2014:ijb.  We
conjecture that this estimator will have improved finite-sample behavior
relative to `iptw_os()` (the alternative estimator described above), though
further study in numerical experiments is warranted.

As with `drtmle()`, the `adaptive_iptw()` function allows missing data in `A`
and `Y`. Similarly, `adaptive_iptw()` also allows for parallelized,
cross-validated estimation of the regression parameters via the `cvFolds` and
`parallel` options.  As with the estimators implemented by `drtmle()`,
cross-validation allows for valid inference under weaker assumptions.

### Outcome-adaptive propensity score estimation

We have recently proposed estimators of the average treatment effect to be used
in settings where this effect is weakly identifiable, as evidenced by estimated
propensity scores of small magnitude.  In these settings, doubly robust
estimators may exhibit non-robust behavior in small samples. The proposed
solution was to target an alternative quantity in the propensity score
estimation step. If we are interested in estimating $\psi_0(a)$ for $a = 0, 1$,
we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$, which
describes the conditional probability of receiving treatment $A = a$ given the
value of the two outcome regressions. For this reason, we call this quantity an
*outcome-adaptive propensity score*. We have shown that all of the usual
asymptotics for TMLE carry through when an estimator of this quantity is
substituted for the true propensity score. The resultant TML estimator is
*super-efficient*, meaning that it will generally have greater asymptotic
efficiency than a standard TML estimator. Simulation experiments indicate that
this approach also delivers more stable behavior in settings where estimated
propensity scores become very small. However, the estimator is not doubly robust
-- the asymptotics rely on the outcome regression being consistently estimated
at $n^{-1/4}$ rate. In this sense, this estimator trades off robustness against
achieving greater stability and efficiency. See @ju2020robust for further
discussion.

#### Implementation of outcome-adaptive propensity score

If the outcome-adaptive PS is desired, the user may specify this via the
`adapt_g` option. In this case, the additional targeting steps needed for doubly
robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed
above, this estimator is *not* doubly robust. Note that when `adapt_g = TRUE`,
the data used for PS estimation will be a `data.frame` with column names `QaW`
where `a` takes all the values specified in `a_0`. For example if `a_0 =
c(0,1)`, then the PS estimation data frame will have two columns called `Q0W`
and `Q1W`. This may be relevant for users writing their own \texttt{SuperLearner}
wrapper functions or those who make use of the `glm_g` setting.

```{r, message = FALSE}
# outcome adaptive propensity score fit using glms
adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
                     adapt_g = TRUE)

# outcome adaptive propensity score fit using super learner
adaptg_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_Q = c("SL.glm", "SL.mean"),
                     SL_g = c("SL.glm", "SL.mean"),
                     adapt_g = TRUE)
```