### 2022-06-25 paper/using_drtmle_obstudies.Rmd
complete

---
output:
  bookdown::pdf_document2:
    template: obs_study_skeleton.tex
    latex_engine: xelatex
bibliography: refs.bib
abs_text: |
  Inverse probability of treatment weighted estimators and doubly robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings, maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \texttt{drtmle} implements doubly robust confidence intervals and hypothesis
  tests for targeted minimum loss estimates of the average treatment effect, in
  addition to several other recently proposed estimators of the average
  treatment effect.
abs_keywords: |
  causal inference, machine learning, semiparametric estimation, treatment
  effects, multiple robustness, targeted minimum loss estimation
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
renv::activate(here::here())
library(nloptr); library(earth)
```

# Introduction

In many fields, researchers are interested in assessing the population-level
effect of a particular treatment on an outcome of interest. The treatment might
correspond to a drug, a potentially harmful exposure, or a policy intervention.
Often, the treatment cannot be randomized due to ethical or logistical reasons.
This has sparked great interest in developing statistical methodology to
estimate population-level effects from observational data. Estimation of these
effects often requires accounting for putative confounding factors -- that is,
factors related to whether a participant receives treatment and to the
participant's outcome. In many settings, researchers measure many potential
confounders, for example using administrative databases or genetic sequencing
technology. Due to the curse of dimensionality, common statistical estimation
techniques are not feasible for such high-dimensional data without restrictive
modeling assumptions. When these assumptions are violated, estimates of the
population-level effect of a treatment can suffer from a large degree of bias.

This has led to increased interest in adopting adaptive estimation techniques
from the machine learning literature to control for high-dimensional confounders
when estimating treatment effects. However, facilitating proper statistical
inference (i.e., confidence intervals and hypothesis tests) about estimates of
effects based on adaptive estimators is challenging. In particular, the use of
standard causal inference techniques, including both G-computation estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa; @robins:2000:epi] is challenging when
utilizing adaptive estimators. In particular, with rare exceptions, if
cross-validation is used to select tuning parameters for the nuisance
parameters, estimates of the population-level effect will be too biased to
achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing* must be
used to select tuning parameters to control the bias [@undersmoothed_hal;
@ertefaie2022nonparametric; @hejazi2022efficient]. However, this can prove quite
challenging in practice, and few guidelines are available.

A more natural framework for integrating adaptive methods into nuisance
regression estimation are doubly robust approaches, including augmented inverse
probability of treatment weighted (AIPTW) estimation [@robins:1994:jasa] and
targeted minimum loss (TML) estimation [@vdlrubin:2006:ijb]. The
asymptotics of these estimators are characterized in part by the convergence
rates of nuisance parameter estimates with respect to a proper norm. This fact
makes the use of cross-validation more appealing for selection of tuning
parameters for estimators of nuisance parameters, thereby providing a more
straightforward means of obtaining an estimate of the effect of interest that
enjoys standard $n^{1/2}$ asymptotics. Based on the efficient influence
function, a key quantity in semiparametric theory, these asymptotic results also
yield closed-form standard error estimates, which can serve as the basis for
constructing confidence intervals and hypothesis tests.

As an intermediate step, the AIPTW and TML estimators require estimation of two
key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and TML
estimators a natural appeal: Inconsistency in the estimation of one nuisance
parameter may be mitigated by the consistent estimation of the other. However,
recent work has shown that the double robustness property does not extend to the
inferential properties of these estimators when adaptive estimators of the PS
and/or OR are used [@vdl:2014:ijb; @benkeser:2017:biometrika]. Instead, valid
inference requires consistent estimation of *both* the PS *and* the OR.
@vdl:2014:ijb proposed new TML estimators that asymptotically attain a normal
sampling distribution with estimable variance under consistent estimation of
*either* the PS *or* the OR, paving the way for inference that is doubly-robust.
@benkeser:2017:biometrika proposed modifications of these estimators with
similar robustness properties and illustrated their practical performance.

The two proposed procedures that yield doubly-robust inference are quite
involved, notably requiring iterative estimation of additional, complex nuisance
parameters. The \texttt{drtmle} package was motivated by the need for a open
source tool to implement these methods. The package implements the TML
estimators of population-level effects proposed in @vdl:2014:ijb and
@benkeser:2017:biometrika, as well as several extensions including
cross-validated targeted minimum loss estimation (CV-TMLE). Additionally,
\texttt{drtmle} implements an IPTW estimator that overcomes shortcomings of
existing IPTW implementations with respect to statistical inference. In
particular, the estimator allows an adaptive estimator of the PS to be used in
construction of the IPTW estimator, while yielding valid statistical inference
about population-level effects. Both the TML and IPTW estimators implemented in
the \texttt{drtmle} package are capable of estimating population-level effects
for discrete-valued treatments; moreover, the package provides convenient
utilities for constructing confidence intervals and hypothesis tests about
contrasts of the mean outcome under different levels of treatment. This
manuscript explains some of the key concepts underlying doubly robust inference
and illustrates usage of the \texttt{drtmle} package for the \texttt{R}
language and environment for statistical computing [@R].

In tandem with growing interest in and usage of doubly robust estimation
approaches, open source software tools implementing these methods have
proliferated. Notably extant software packages include the \texttt{tmle}
\texttt{R} package [@gruber2012tmle], which implements TML estimators of the
effects of static single time point interventions; the \texttt{ltmle} \texttt{R}
package [@lendle2017ltmle], which implements TML estimators of the effects of
multiple time point interventions, subject to time-dependent confounding and
censoring, as well as estimators of the parameters of a longitudinal working
marginal structural model; the \texttt{tmle3} \texttt{R} package
[@coyle2021tmle3-rpkg], which serves as a generalizable engine for implementing
TML estimators within the \texttt{tlverse} ecosystem [@vdl2022targeted],
designed to provide a common and extensible framework for TML estimation of the
effects of static, dynamic, and stochastic single time point interventions; and
the \texttt{doubleml} \texttt{R} and \texttt{Python} packages
[@bach2021doublemlR; @bach2022doublemlPy], which implement estimators of causal
effects within the double machine learning framework [@chernozhukov2017double;
@chernozhukov2018double], reyling upon Neymon orthogonality and sample-splitting
for robustness. From among this non-exhaustive list, \texttt{drtmle} stands out
as, to our knowledge, the only implementation of statistical procedures designed
to attain doubly robust inference [@benkeser:2017:biometrika], uniquely
positioning \texttt{drtmle} as a robust tool for the estimation of treatment
effects in observational data settings.

# Definition and estimators of the average treatment effect

## Parameters of interest

Suppose the observed data consist of $n$ independent and identically distributed
copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is a vector of
baseline covariates, $A \in \{0,1\}$ is a binary treatment, $Y$ is an outcome,
and $P_0$ is the true data-generating distribution. The parameters of interest
are
\begin{equation}
  \psi_0(a) = E_0\{\overline{Q}(a,W)\} = \int \overline{Q}_0(a,w) dQ_{W,0}(w)
  \ , \ \mbox{for} \ a = 0,1 \ ,
  \label{parameterDef}
\end{equation}
where $\overline{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
assumptions, these parameters have interpretations as mean counterfactual
outcomes under the different treatments, with contrasts between these means
defining causal effects.

## G-computation estimators

An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
Suppose we have available an estimate of the OR, $\overline{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
$\psi_n(a) = \int \overline{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n
  \overline{Q}_n(a, W_i) \ .
  \label{eqn:gcomp}
\end{equation}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

## Inverse probability of treatment estimators

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
\begin{align}
  E_0\{\overline{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
  \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
(\ref{iptwID}) motivates the estimator
\begin{equation*}
  \psi_{n,\text{IPTW}}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap;
however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
are constructed as solutions to an IPTW estimating equation, the empirical
variance of which may also be used for standard error estimation. If adaptive
approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing, as explored by
@ertefaie2022nonparametric and @hejazi2022efficient.

## Doubly robust estimators

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness -- that is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators generally lack a valid basis for asymptotic statistical inference.

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One prominent example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\}
  \overline{Q}_n(a, W_i) \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators: It exhibits _double robustness_ in that it is consistent for
$\psi_0(a)$ if either $\overline{Q}_n$ is consistent for the true OR *or* $g_n$ is
consistent for the true PS. Furthermore, if both the OR and PS are consistently
estimated and regularity conditions hold, one can establish a normal limiting
distribution for the AIPTW estimator. Closed-form estimators of the variance of
this limiting distribution can be derived that may be used to construct
Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss estimation (or TMLE).
TMLE is a general methodology that may be used to modify an estimator of a
regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\overline{Q}_n^{\star}$ satisfies
\begin{equation}\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n^{\star}(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\overline{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies (\ref{oneStepEst}). In addition to possessing the
desirable asymptotic properties, TML estimators have been shown to outperform
AIPTW estimators in finite samples [@porter:2011:ijb]. We refer readers to
@tlbook:2011 and @vdl2022targeted for further details about on the TMLE
methodology.

# Doubly robust inference

@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. @vdl:2014:ijb derived
modified TML estimators that are doubly robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly robust variance estimators that may be used to
construct doubly robust confidence intervals and hypothesis tests.
@benkeser:2017:biometrika proposed simplifications of @vdl:2014:ijb's
estimators and demonstrated the practical performance of these estimators via
simulation experiments.

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (\ref{oneStepEst}) in addition to
two additional equations. These equations involve additional regression
parameters that we refer to as the reduced outcome regression (R-OR) and the
reduced propensity score (R-PS), defined respectively as
\begin{align}
  \overline{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \overline{Q}_n(W) \mid
  A = a, g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qoverliner} \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w), g_n(W)=g_n(w)\right\} \ .
  \label{eqn:gr_bivariate}
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @vdl:2014:ijb showed that if estimates of the
OR $\overline{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR
$\overline{Q}_{r,n}^{\star}$, and R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \overline{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\overline{Q}_{r,n}(a, W_i)}
{g_n(a \mid W_i)}\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \overline{Q}_n^{\star}(a, W_i) \}
= 0 \ ,
\label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\overline{Q}_n^{\star}$ would be doubly
robust not only with respect to consistency, but also with respect to its
limiting distribution. The author additionally proposed a TMLE algorithm to map
initial estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy
these key equations.

@benkeser:2017:biometrika demonstrated alternative equations that can be
satisfied to achieve this form of double-robustness. In particular, these
authors formulated an alternative version of the R-PS,
\begin{align*}
  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w) \} \ ,
\end{align*}
which we refer to as R-PS1. They also introduced an additional regression of a
weighted residual from the PS on the OR,
\begin{align*}
  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
  {g_n(a \mid W)} \ \bigg\vert \ \overline{Q}_n(W) = \overline{Q}_n(w) \biggr\} \ ,
\end{align*}
which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
both univariate regressions, while R-PS is a bivariate regression.
@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
easier to estimate consistently than R-PS. Further, they showed that if
estimators of the OR, PS, and R-OR satisfy equations
(\ref{meanEIFZero})--(\ref{eqn:qoverliner}), and if, additionally, estimators of
the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2 $g_{r,n,2}^{\star}$ satisfy
\begin{equation*}
\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
\overline{Q}_n^{\star}(a,W_i)\} = 0 \ ,
\end{equation*}
then the GCOMP estimator based on this $\overline{Q}_n^{\star}$ also enjoys a
doubly robust limiting distribution. @benkeser:2017:biometrika provided a
TMLE algorithm that produced such estimates and provided closed-form,
doubly robust variance estimates.

\textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
methods for constructing doubly robust (with respect to consistency) estimators
of our parameter of interest. Theoretically, these two frameworks are closely
related and, practically, the two estimators are generally seen as competitors
for estimating marginal means. In equation (\ref{oneStepEst}), we were able to
generate a doubly robust estimator by adding a term to the GCOMP estimator. This
term is exactly the left-hand side of the key equation (\ref{meanEIFZero}), the
key equation
that is solved by the standard TMLE. This is no coincidence -- this term is
crucial in studying the asymptotic properties of the GCOMP, AIPTW, and TML
estimators (for discussion, see @benkeser:2017:biometrika, Section 2.2). Recall
that a TML estimator achieves doubly-robust consistency by satisfying equation
(\ref{meanEIFZero}), while the AIPTW achieves doubly robust consistency by
combining the GCOMP estimator with the left-hand side of this term. Now, we have
argued that a TML estimator may achieve doubly robust limiting behavior if it
additionally satisfies equations (\ref{eqn:qoverliner}) and
(\ref{eqn:gr_bivariate}). It is thus sensible to wonder whether an AIPTW-style
estimator could be constructed by combining the left-hand side of these two
additional equations with the AIPTW estimator. @benkeser:2017:biometrika showed
that, surprisingly, this AIPTW-style estimator does not achieve the same
doubly robust properties as the TML estimator. Thus, while AIPTW and TMLE are
generally competitors for producing estimators doubly robust with respect to
consistency, for doubly robust inference, TML estimators are preferable.

# Implementation of doubly robust inference

The main workhorse of the \texttt{drtmle} package is the eponymous `drtmle()`
function. This function estimates the treatment-specific marginal mean for
user-specified levels of a discrete-valued treatment and computes a doubly
robust covariance matrix for these estimates. The function implements either the
estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the improved
estimators of @benkeser:2017:biometrika (if `reduction = "univariate"`). The
package includes utility functions for combining these estimates into estimates
of average treatment effects, as well as other contrasts, as discussed below.

The main data arguments for `drtmle()` are `W, A, Y`, which, as above, represent
the covariates, a discrete-valued treatment, and an outcome, respectively.
Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
and R-PS (or R-PS1 and R-PS2). The package provides flexibility in how these
regression parameters may be estimated. Once the user has specified the data
arguments and how to estimate each regression parameter, the function proceeds
as follows:
  1. estimate the OR via the internal `estimateQ()` function;
  2. estimate the PS via the internal `estimateG()` function;
  3. estimate reduced-dimension regression via the internal `estimateQrn()` and
     `estimategrn()` functions;
  4. iteratively modify initial estimates of the OR and PS via the
     `fluctuateQ()` and `fluctuateG()` functions until equations
     (\ref{meanEIFZero})--(\ref{eqn:gr_bivariate}) are approximately solved;
  5. compute the TML estimates and covariance estimates based on the final OR.
We refer interested readers to @benkeser:2017:biometrika for a theoretical
derivation of how the iterative modification of initial OR and PS is performed.

In addition to returning doubly robust parameter and covariance estimates, the
`drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
modified AIPTW estimator discussed in the remark above. While doubly robust
inference based on these estimators is not available, the estimators are
provided for comparison. In the following subsections, we demonstrate the use
of these methods by examining various options for making calls to `drtmle()`.

## Estimating regressions
```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
```
While the estimators computed by `drtmle()` are consistent and asymptotically
normal under inconsistent estimation of either the OR or PS, it is nevertheless
advisable to strive for consistent estimation of *both* the OR and PS. If both
of the regression parameters are consistently estimated, then the estimators
computed by `drtmle()` achieve the nonparametric efficiency bound (that is,
the asymptotic variance of the efficient influence function), resulting in
narrower confidence intervals and more powerful hypothesis tests. In order for
the estimates of the OR and PS to have the greatest chance of consistency, the
`drtmle()` function facilitates estimation using adaptive estimation techniques.
In fact, there are three ways to estimate each of the regression parameters:
generalized linear models, super learning [@vdl:2007:statapp;
@SuperLearnerPackage], and a user-specified estimation technique. We demonstrate
each of these approaches in turn on a simulated data set in which the treatment
is not randomized. The true parameter values of interest are $\psi_0(0)=$ `r
round(EY0, 2)` and $\psi_0(1)=$ `r round(EY1, 2)`.

```{r}
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 1, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
```

### Using generalized linear models to estimate regressions

Generalized linear models are perhaps the most popular means of estimating
regression functions. On account of this legacy, these estimators have been
included as an option for estimating the regression parameters. The user
specifies the right-hand side of a regression formula as a character string
in the `glm_g`, `glm_Q`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS,
R-OR, and R-PS. These regressions are fit by calling the `glm()` function from
\texttt{R}'s base \texttt{stats} package [@R]. Thus, the model specification
should be able to be parsed as a valid formula object, as required by `glm()`.
For the PS formula specified by `glm_g`, `colnames(W)` are available to use in the
formula. For the OR (`glm_Q`) available names of variables to be used in the
formula depends on the `stratify` option. If `stratify = FALSE`, a single OR is
fit to estimate $\overline{Q}_0(A,W)$. In this case, the OR regression is fit by
passing a `data.frame` that contains `W` and `A` into the `data` argument of
`glm()`. Thus, both `colnames(W)` and `"A"` may appear in the formula specified
in `glm_Q`. If `stratify = TRUE`, regressions are fit stratifying on `A` to
separately estimate $\overline{Q}_0(1,W)$ and $\overline{Q}_0(0,W)$. In this
case, the `glm_Q` formula may not include `"A"`.

To understand the variable names available for R-OR, recall from equation
(\ref{eqn:qoverliner}) that the nuisance parameter involves a regression onto
the estimated propensity score. Thus, the formula that is input via `glm_Qr`
should assume data with a single column, `"gn"`.  Similarly, the R-PS1 and R-PS2
that are computed for the estimators of @benkeser:2017:biometrika (the default
estimators, computed whenever `reduction = "univariate"`) involve regressions
onto the estimated outcome regression. Thus, the formula `glm_gr` should assume
data with a single column named `"Qn"` if `reduction = "univariate"`. Instead,
if the estimators of @vdl:2014:ijb are preferred (`reduction = "bivariate"`),
the `glm_gr` formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes
data with two columns, `"Qn"` and `"gn"`.

The code below illustrates how generalized linear models can be used to
estimate the regression parameters. In this call to `drtmle()`, we specify
`family = binomial()` so that logistic regression is used to estimate the OR.
We illustrate non-stratified regression for both the univariate (default) and
bivariate reductions below.

```{r}
glm_fit_uni <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE)
glm_fit_uni

glm_fit_biv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                      glm_gr = "Qn + gn", glm_Qr = "gn", stratify = FALSE,
                      reduction = "bivariate")
glm_fit_biv
```

### Using super learner to estimate regressions

```{r, echo = FALSE, message = FALSE}
library(SuperLearner)
```
Super learning is a loss-based ensemble machine learning technique
[@vdl:2007:statapp] that generalizes stacking and stacked regression
[@wolpert:1992:nnet; @breiman:1996:ml]. The user specifies many potential
candidate regression algorithms, referred to as a *library* of candidate
estimators, and the super learner algorithm estimates the convex combination of
regression estimators that minimizes the $V$-fold cross-validated risk based on
a user-selected loss function [@dudoit2005asymptotics]. Oracle inequalities
show that the super learner estimate performs essentially as well as the unknown
best convex combination of the candidate estimators [@vdvaart:2006:statdec].
Thus, the methodology may be seen as an asymptotically optimal way of performing
estimator selection. Practically, the super learner provides the best chance
for obtaining a consistent estimator for both the OR and PS, and thereby an
efficient estimator of the marginal means of interest.

Super learning is implemented in the R package \texttt{SuperLearner}
[@SuperLearnerPackage] and `drtmle()` provides the option to utilize super
learning for estimation of the OR, PS, and reduced-dimension regressions via the
`SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle()` either the
generalized linear model options for regression modeling or the super learner
options should be invoked. If both are called, the function will automatically
opt to using the super learner option. The inputs to the super learner options
are passed to the `SL.library` argument of the `SuperLearner()` function of the
\texttt{SuperLearner} package. We refer readers to the documentation of the
`SuperLearner()` function for information on how to properly specify a super
learner library. In the code below, we illustrate a call to `drtmle()` with a
relatively simple library. For the OR and PS, the library includes a main-terms
generalized linear model (via the function `SL.glm` from \texttt{SuperLearner}) and
an unadjusted generalized linear model (`SL.mean`). For the reduced-dimension
regressions we use a super learner library with a main-terms generalized linear
model (`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
\texttt{earth} package [@earthPackage]).

```{r, warning = FALSE, message = FALSE, cache = TRUE}
set.seed(1234)
sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                 SL_g = c("SL.glm","SL.mean"),
                 SL_Q = c("SL.glm","SL.mean"),
                 SL_gr = c("SL.glm", "SL.earth"),
                 SL_Qr = c("SL.glm", "SL.earth"),
                 stratify = FALSE)
sl_fit
```

The \texttt{drtmle} package also allows users to pass in their own functions for
estimating regression parameters via the `SL_` options. These functions must be
formatted properly for compatibility with `drtmle()`. The necessary formatting
is identical to the formatting required for writing a `SuperLearner` wrapper
(see `SuperLearner::write.SL.template()`). The \texttt{drtmle} package includes
`SL.npreg`, which is an example of how a user can specify a function for
estimating regression parameters. This functions fits a kernel regression
estimator using the \texttt{np} package [@npPackage]. Below we show the source
code for this function.

```{r}
SL.npreg
```

In the above code, we see that a user-specified function should have options
`Y, X, newX, family, obsWeights, ...`, along with other options specific to the
function. The option `Y` corresponds to the outcome of the regression and `X`
to the variables onto which the outcome is regressed. The option `newX` should
be of the same class and dimension as `X` and is meant to contain new values of
variables at which to evaluate the regression fit. The `family` and
`obsWeights` options are not strictly necessary (indeed, they are not used by
`SL.npreg`), though these may be useful to include if one wishes to utilize the
function as part of a super learner library. The `SL.npreg` function proceeds by
checking whether there is sufficient variation in `Y` to justify fitting the
kernel regression (as specified by option `rangeThresh`), and if so, fits a
kernel regression with cross-validated bandwidth selection via the `npregbw()`
function. The output is then formatted in a specific way so that `drtmle()` is
able to retrieve the appropriate objects from the fitted regression. In
particular the output should be a list with named objects `pred` and `fit`; the
former including predictions made from the fitted regression on `newX`, the
latter including any information that may be needed to predict new values based
on the fitted object. A class is assigned to the `fit` object, so that an S3
`predict()` method may later be used to predict new values based on the fitted
regression object. The user must also specify this `predict()` method. Use
`drtmle:::predict.SL.npreg()` to view the `predict()` method for
`SL.npreg`.

In the code below, we demonstrate a call to `drtmle()` with `SL.npreg` used to
estimate each nuisance parameter. Here, we also illustrate the use of `stratify
= TRUE`, which fits a separate kernel regression of $Y$ onto $W$ in
observations with $A = 1$ and $A = 0$.

```{r, cache = TRUE, message = FALSE}
npreg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = "SL.npreg", SL_Q = "SL.npreg",
                    SL_gr = "SL.npreg", SL_Qr = "SL.npreg",
                    stratify = TRUE)
```

In the Appendix, we describe an additional means of using `drtmle()` with
OR and PS estimates that have been obtained externally.

### Reducing dependence on random seed

Some users may find it unsettling that when applying the super learner in real
data analysis the results are dependent on the random seed that is set. While
this is a natural artifact of cross-validation (sample-splitting) and adaptive
estimation, one can average results over repeated super learner runs to reduce
or remove this dependence -- more repeats will lead to less dependence on the
random seed. There are generally two approaches that can be used: (i) averaging
on the scale of the nuisance parameters and (ii) averaging on the scale of the
point estimates.

For (i), suppose that we have $n_{SL}$ estimated super learners for the OR,
$\overline{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS $g_{n,j}, j = 1, \dots,
n_{SL}$. We can take the average over these fits,
\begin{equation*}
  \overline{Q}_n = \frac{1}{n_{SL}} \sum_{j = 1}^{n_{SL}} \overline{Q}_{n,j}
\end{equation*}
as our estimate of the OR (similarly for the PS). These averaged estimates are
then used to compute the AIPTW or TMLE, as described above.

For (ii), suppose for example that we have $n_{SL}$ TML (similarly, AIPTW)
estimates $\psi_{n,j}(1)$ of $\psi_0(1)$, each obtained based on one (or more)
super learners of the OR and/or PS. We can take the average over these point
estimates,
\begin{equation*}
  \psi_n(1) = \frac{1}{n_{SL}} \sum_{j=1}^{n_{SL}} \psi_{n,j}(1)
\end{equation*}
as our estimate of $\psi_{0}(1)$.

This feature is implemented via the `n_SL` and `avg_over` options, where the
former specifies the number of super learners to repeat and the latter
specifies the scale(s) over which to average. For example, if `avg_over =
"drtmle"` and `n_SL = 2`, then two point estimates are computed, each based on a
single super learner, and averaged to obtain a final estimate. If instead
`avg_over = "SL"` and `n_SL = 2`, then two super learners are fit and averaged
before a final point estimate is obtained. Finally, if `avg_over = c("drtmle",
"SL")` and `n_SL = 2` then averaging on *both* scales is performed. That is, the
final estimate is the average of two point estimates, each obtained using OR and
PS estimates that are averaged over two super learners.

Unsurprisingly, averaging over multiple super learners can add considerable
computational expense to calls to `drtmle()`. The package is currently
parallelized at the level of individual calls to `SuperLearner()`. With
parallelization over enough cores, repeated super learning should not add too
much to the computational burden relative to a single call to `SuperLearner()`.

## Standard error estimation

This section contains some more advanced mathematical material and may be
skipped over by users interested only in implementation. For simplicity, we
focus this discussion on the standard TML and AIPTW estimators. The ideas
generalize immediately to the DR-TML estimator, but the formulas are more
complex in this latter case.

Reported standard error estimates are based on an estimate of the asymptotic
variance of the estimators. This variance can be characterized by the
(efficient) influence function of the estimators. For example, let
$\psi_{n}(1)$ be an AIPTW estimate of $\psi_0(1)$. Under regularity conditions,
\begin{equation*}
  \psi_{n}(1) - \psi_0(1) = \frac{1}{n} \sum_{i=1}^n
    D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
\end{equation*}
where
\begin{equation*}
  D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}
  {g_0(1 \mid W_i)} \{ Y_i - \overline{Q}_0(1, W_i)\} + \overline{Q}_0(1, W_i)
  - \int \overline{Q}_0(1, w) dQ_{W,0}(w) \ .
\end{equation*}
Thus, by the central limit theorem, $n^{1/2} \psi_{n}(1)$ converges in
distribution to a normal distribution with mean $\psi_0(1)$ and variance
\begin{equation*}
  \sigma^2_0 = E_0( [ D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O) -
  E_0\{D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
\end{equation*}
An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical
variance of $D_i = D^{\star}(\overline{Q}_n, g_n, Q_{W,n})(O_i)$, that is,
\begin{equation*}
  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{D_i - \frac{1}{n} \sum_{j=1}^n
  D_j \right\}^2 \ .
\end{equation*}
Thus, an estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n /
n^{1/2}$. Conveniently, this idea generalizes naturally to estimating the
asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.

### Standard errors for TMLE

For TML estimators, it is standard practice to construct $D_i$ using the
*targeted* version of the OR (denoted by $\overline{Q}_n^{\star}$ above). Recall that
such estimates are obtained by modifying an initial OR estimate so as to ensure
a particular (set of) equation(s) is (are) solved.  However, simulations have
shown that better standard error estimates for TMLE may be achieved by instead
using the *initial* OR estimate in the computation of standard errors, as
described in the preceding subsection. To this end, The `drtmle()` function
includes the `targeted_se` option, which specifies whether initial nuisance
parameter estimates are to be in computing the standard error (i.e.,
`targeted_se = FALSE`).

```{r, show_targeted_se, cache = TRUE}
glm_fit_uni_nontargeted_se <-
  drtmle(W = W, A = A, Y = Y, family = binomial(),
         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
         glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE,
         targeted_se = FALSE)
# compare to original call
list(
  glm_fit_uni$drtmle$cov, # based on targeted OR
  glm_fit_uni_nontargeted_se$drtmle$cov # untargeted OR
)
```

### Cross-validated standard errors

Simulation studies have revealed that when employing machine learning to
estimate the OR and/or PS, the estimated standard errors are often *too small*,
in small to moderate-sized samples. Problematically, this leads directly to
under-coverage of confidence intervals and possibly inflated Type-I error rates
of hypothesis tests based on these standard error estimates. This should not be
too surprising, as the phenomenon is related to the form of the standard error
estimates. In particular, such estimates include a term that resembles a
(weighted) mean squared error (MSE) of the OR. Since machine learning algorithms
often set out to minimize empirical MSE as part of their training process, it
follows that the empirical MSE often ends up biased towards zero relative to the
true MSE. To more accurately estimate the MSE of a machine learning algorithm,
cross-validation can be used. The same sample-splitting idea can be used to
obtain more conservative standard error estimates.

Suppose we use $V$-fold cross-validation to obtain $V$ training-sample-specific
estimates of the OR, $\overline{Q}_{n,v}, v = 1, \dots, V$ and the PS,
$g_{n,v}, v = 1, \dots, V$. For $i = 1, \dots, n$, let
$\overline{Q}_{n,i}^{cv}(1)$ denote the estimate of $\overline{Q}_{0}(a, W_i)$
obtained when observation $i$ was in the validation fold. That is,
$\overline{Q}_{n,i}^{cv}$ is an out-of-sample prediction of $Y_i$. We define
$g_{n,i}^{cv}$ similarly. Using these estimates, we can define
\begin{equation*}
 D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i -
 \overline{Q}_{n,i}^{cv}(1)\} + \overline{Q}_{n,i}^{cv}(1) -
 \frac{1}{n} \sum_{i=1}^n \overline{Q}_{n,i}^{cv}(1)
\end{equation*}
and compute an estimate of $\sigma^2_0$ using
\begin{equation*}
  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} -
  \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ .
\end{equation*}

In \texttt{drtmle}, cross-validated standard error estimates can be obtained by
setting `se_cv = "full"` and `se_cvFolds` to a number greater than one. For
example,

```{r, fit_w_full_cv, warning = FALSE, cache = TRUE}
glm_fit_uni_cv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         se_cv = "full", se_cvFolds = 2, stratify = FALSE,
                         targeted_se = FALSE) # needed for this to work
# compare variance to previously obtained
list(
  `No CV` = glm_fit_uni$drtmle$cov,
  `CV` = glm_fit_uni_cv$drtmle$cov
)
```

### Partially cross-validated standard errors

Obtaining fully cross-validated standard errors can be fairly time-intensive
when super learner is used to estimate the OR and PS. In \texttt{drtmle}, we
give an option to compute standard error estimates that are *partially
cross-validated* and do not require any additional fitting beyond a standard
super learner. Here, we describe briefly how these estimates are obtained and
show calls to `drtmle()` for obtaining such estimates.

Suppose we have $K$ candidate regressions in our super learner library for both
the OR and PS (in practice, the number could be different for the OR and for the
PS). Recall that a super learner estimate of, for example, $\overline{Q}_0(a,
w)$ is obtained as a weighted combination of estimates from each candidate
regression. Let $\overline{Q}_{n,k}$ denote the estimate of $\overline{Q}_{0}$
obtained by training algorithm $k$ using the full data set. The super learner
estimate of $\overline{Q}_{0}(a,w)$ is
\begin{equation*}
  \overline{Q}_{n,SL}(a, w) = \sum_{k = 1}^K \alpha_{n,k}
  \overline{Q}_{n,k}(a, w) \ ,
\end{equation*}
where $\alpha_n = (\alpha_{n,1}, \dots, \alpha_{n,K})$ is a weight vector that
is selected by minimizing a cross-validated risk criterion. That is, each of the
$K$ algorithms is trained according to a $V$-fold cross-validation scheme,
resulting in $V + 1$ fits of each algorithm (one in each of $V$ training folds,
plus one obtained by fitting the algorithm using the full data). Let
$\overline{Q}_{n,k,v}$ denote the $k$\textsuperscript{th} algorithm trained
using the $v$\textsuperscript{th} training fold. For $i = 1, \dots, n$, let
$\overline{Q}_{n,k,i}^{cv}(1)$ denote the estimate of $\overline{Q}_0(1, W_i)$
implied by the fitted algorithm $\overline{Q}_{n,k,v}$, which was obtained when
observation $i$ was in the validation set. That is, for observations with
$A_i = 1$, $\overline{Q}_{n,k,i}^{cv}(1)$ is an out-of-sample prediction of
$Y_i$ based on the $k$\textsuperscript{th} algorithm. Now, we can construct a
*partially cross-validated* super learner prediction as follows. For
$i = 1, \dots, n$ let
\begin{equation*}
  \overline{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k}
    \overline{Q}_{n,k,i}^{cv}(1)
\end{equation*}
be the estimate obtained by ensembling the training-set-specific algorithm fits
based on $\alpha_n$. We can use these estimates to produce our *partially
cross-validated* standard error estimate as follows. First, define
\begin{equation*}
  D_i^{pcv} = \frac{I(A_i = 1)}{g_{n,SL,i}^{cv}} \{ Y_i -
    \overline{Q}_{n,SL,i}^{cv}(1)\} + \overline{Q}_{n,SL,i}^{cv}(1) -
    \frac{1}{n} \sum_{i=1}^n \overline{Q}_{n,SL,i}^{cv}(1),
\end{equation*}
then, compute an estimate of $\sigma^2_0$ using
\begin{equation*}
  \sigma^2_{n,pcv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{pcv} -
  \frac{1}{n} \sum_{j=1}^n D_j^{pcv} \right\}^2 \ .
\end{equation*}

Note that is not a properly cross-validated estimate, since observation $i$ is
used to compute the weight vector $\alpha_n$. However, such estimates can be
obtained based on the output of a single round of super learning, thereby
providing improved standard error estimates at minimal additional computational
cost.

Below, we demonstrate how to obtain such estimates using `drtmle()`.

```{r, demo_pcv, cache = TRUE, warning = FALSE}
set.seed(1234)
sl_fit_pcv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm","SL.mean"),
                     SL_Q = c("SL.glm","SL.mean"),
                     SL_gr = c("SL.glm", "SL.earth"),
                     SL_Qr = c("SL.glm", "SL.earth"),
                     stratify = FALSE,
                     se_cv = "partial")
# compare estimates
# pt estimates should be same
# variance should be larger for pcv
list(
  `No-CV` = sl_fit$drtmle$cov, # no cv
  `Partial CV` = sl_fit_pcv$drtmle$cov # partial cv
)
```

## Confidence intervals

The \texttt{drtmle} package contains methods that compute doubly robust
confidence intervals. The `ci()` method computes confidence intervals about the
estimated marginal means in a fitted object of S3 class `"drtmle"`, in addition
to contrasts between those means. By default, the method gives confidence
intervals about each mean.

```{r}
ci(npreg_fit)
```

Alternatively the `contrast` option allows users to specify a linear
combination of marginal means, which can be used to estimate the average
treatment effect, as we show below.

```{r}
ci(npreg_fit, contrast = c(1, -1))
```

The `contrast` option also accepts as input a list of functions to be used to
compute a confidence interval for a user-specified contrast. For example, we
might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
constructing Wald-style confidence intervals for ratios, it is common to
compute the interval on the logarithmic scale and subsequently back-transform.
That is, the confidence interval takes the form
\begin{equation*}
  \exp \biggl[\log \biggl\{ \frac{\psi_n(1)}{\psi_n(0)} \biggr\} \pm
  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
\end{equation*}
where $\sigma^{\log}_n$ is the estimated standard error on the logarthmic scale.
By the delta method, an estimate of this standard error is given by
\begin{equation*}
  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
\end{equation*}
where $\Sigma_n$ is the doubly robust covariance matrix estimate output from
`drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
This confidence interval may be obtained as follows.

```{r}
riskRatio <- list(f = function(eff){ log(eff) },
                  f_inv = function(eff){ exp(eff) },
                  h = function(est){ est[1] / est[2] },
                  fh_grad =  function(est){ c(1 / est[1], -1 / est[2]) })
ci(npreg_fit, contrast = riskRatio)
```

More generally, this manner of inputting the `contrast` option to the `ci()`
method can be used to compute confidence intervals of the form
\begin{equation*}
  f^{-1}\bigl\{ f(h(\psi_n)) \pm z_{1-\alpha/2} \nabla f(h(\psi_n))^T \Sigma_n
  \nabla f(h(\psi_n)) \bigr\} \ ,
\end{equation*}
where $f$ (`contrast$f()`) is the transformation of the confidence interval,
$f^{-1}$ (`contrast$f_inv()`) is the back-transformation of the interval, $h$
(`contrast$h()`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
(`contrast$fh_grad()`) defines the gradient of the transformed contrast at the
estimated marginal means.

We note that this method of computing confidence intervals can also be used to
obtain a transformed confidence interval for a particular marginal mean. For
example, recall that $Y$ is binary in our running example. Thus, we may wish to
enforce that our confidence interval be computed on a scale ensuring that the
estimated confidence limits fall within the unit interval $[0, 1]$. To that end,
we can utilize an interval of the form:
\begin{equation*}
  \text{expit} \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
  \pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
\end{equation*}
This confidence interval may be implemented as follows.

```{r}
logitMean <- list(f = function(eff){ qlogis(eff) },
                  f_inv = function(eff){ plogis(eff) },
                  h = function(est){ est[1] },
                  fh_grad = function(est){ c(1 / (est[1] - est[1]^2), 0) })
ci(npreg_fit, contrast = logitMean)
```

## Hypothesis tests

Doubly robust Wald-style hypothesis tests may be implemented using the
`wald_test()` method. As with confidence intervals, there are three options for
testing. First, we may test the null hypothesis that the marginal means equal a
fixed value (or values). Below, we test the null hypothesis that $\psi_0(1) =
0.5$ and that $\psi_0(0) = 0.6$ by inputting a vector to the `null` option.

```{r}
wald_test(npreg_fit, null = c(0.5, 0.6))
```

As with the `ci()` method, the `wald_test()` method allows users to test linear
combinations of marginal means by inputting a vector of ones and negative ones
in the `contrast` option. We can use this to test the null hypothesis of no
average treatment effect or to test the null hypothesis that the average
treatment effect equals a given reference value, e.g., $0.05$, by specifying
the `null` option.

```{r}
wald_test(npreg_fit, contrast = c(1, -1))
wald_test(npreg_fit, contrast = c(1, -1), null = 0.05)
```

The `wald_test()` method also allows for user-specified contrasts to be tested
using syntax similar to the `ci()` method. The only difference syntactically is
that it is not strictly necessary to specify `f_inv()`, as the hypothesis test is
performed on the transformed scale. That is, we can generally test the null
hypothesis that $f(h(\psi_0))$ equals $f_0$ (the function $f$ applied to the
value passed to `null`) using the Wald statistic
\begin{equation*}
  Z_n \coloneqq \frac{\{f(h(\psi_n)) - f_0\}}{\nabla f(h(\psi_n))^T \Sigma_n
  \nabla f(h(\psi_n))} \ .
\end{equation*}
We can use the `riskRatio` contrast defined previously to test the null
hypothesis that $\psi_0(1) / \psi_0(0) = 1$.

```{r}
wald_test(npreg_fit, contrast = riskRatio, null = 1)
```

## Missing data

The `drtmle()` function supports missing data in `A` and `Y`. Such missing data
are common in practice and, if ignored, can bias estimates of the marginal
means of interest. The estimators previously discussed can be modified to allow
such missingness processes to depend on observed covariates. Let $\Delta_A$ and
$\Delta_Y$ be indicators that $A$ and $Y$ are observed, respectively. We can
redefine the OR to be $\overline{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
A = a, \Delta_A = 1, \Delta_Y = 1, W = w)$ and similarly redefine the PS to be
\begin{align}
  g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
  \nonumber \\
  &= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
  \label{modPS} \\
  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w)
   \nonumber.
\end{align}

To demonstrate, we artificially introduce missing values to `A` and `Y` into
the data from our running example.

```{r}
DeltaA <- rbinom(n, 1, plogis(2 + W$W1))
DeltaY <- rbinom(n, 1, plogis(2 + W$W2 + A))
A[DeltaA == 0] <- NA
Y[DeltaY == 0] <- NA
```

The only required modification of the call to `drtmle()` is in how the PS
estimation is performed. The options `glm_g` and `SL_g` are still used to fit
generalized linear models or a super learner, respectively. However, the code
allows for separate regression fits for each of the three components of the PS
in equation (\ref{modPS}). To perform separate generalized linear model fits for
each of the three components, `glm_g` should be a named list with entries
`"DeltaA"`, `"A"`, and `"DeltaY"`. Respectively, these should provide a formula
expression for the regression of $\Delta_A$ on $W$, $A$ on $W$ in observations
with $\Delta_A = 1$, and of $\Delta_Y$ on $W$ and $A$ in observations with
$\Delta_A = \Delta_Y = 1$ (if `stratify = FALSE`) _or_ of $\Delta_Y$ on $W$ to
be fit in observations with $\Delta_A = \Delta_Y = 1$ and $A = a$ for each $a$
specified in option `a_0` (if `stratify = TRUE`). If only a single formula is
passed to `glm_g`, it is used for each of the three regressions. We illustrate a
call to `drtmle()` with missing data and generalized linear model fits below.

```{r, cache = TRUE}
glm_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                       glm_g = list(DeltaA = "W1 + W2", A = "W1 + W2",
                                    DeltaY = "W1 + W2 + A"),
                       glm_Q = "W1 + W2*A", glm_Qr = "gn",
                       glm_gr = "Qn", family = binomial())
glm_fit_wNAs
```

It is possible to mix generalized linear models and super learning when fitting
each of the components of the PS. In the example below, we use the wrapper
function `SL.glm`, which fits a main terms logistic regression, to fit the
regression of $\Delta_A$ on $W$; use `SL.npreg` to fit the regression of $A$ on
$W$; and use a super learner with library `SL.glm`, `SL.mean`, and `SL.earth` to
fit the regression of $\Delta_Y$ on $W$ and $A$.

```{r, cache = TRUE}
mixed_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         SL_g = list(DeltaA = "SL.glm", A = "SL.npreg",
                         DeltaY = c("SL.glm", "SL.mean", "SL.earth")),
                         glm_Q = "W1 + W2*A", glm_Qr = "gn",
                         glm_gr = "Qn", family = binomial())
```

We can also fit the PS regressions externally and pass in via `gn`, though this
process is slightly more complicated than when no missing data are present. The
basic idea is to fit one regression for (i) the indicator of missing $A$, (ii)
$A$ itself, and (iii) the indicator of the outcome being missing. The three PS's
are then multiplied together and an appropriately ordered list is constructed as
before. We illustrate with the following example.

```{r, cache = TRUE}
# first regress indicator of missing A on W
fit_DeltaA <- glm(DeltaA ~ . , data = W, family = binomial())

# get estimated propensity for observing A
ps_DeltaA <- predict(fit_DeltaA, type = "response")

# now regress A on W | DeltaA = 1
fit_A <- glm(A[DeltaA == 1] ~ . , data = W[DeltaA == 1, , drop = FALSE],
             family = binomial())

# get estimated propensity for receiving A = 1
ps_A1 <- predict(fit_A, newdata = W, type = "response")
# propensity for receiving A = 0
ps_A0 <- 1 - ps_A1

# now regress DeltaY on A + W | DeltaA = 1. Here we are pooling over
# values of A (i.e., this is what drtmle does if stratify = FALSE).
# We could also fit two regressions, one of DeltaY ~ W | DeltaA = 1, A = 1
# and one of DeltaY ~ W | DeltaA = 1, A = 0 (this is what drtmle does if
# stratify = TRUE).
fit_DeltaY <- glm(DeltaY[DeltaA == 1] ~ . ,
                  data = data.frame(A = A, W)[DeltaA == 1, ],
                  family = binomial())

# get estimated propensity for observing outcome if A = 1
ps_DeltaY_A1 <- predict(fit_DeltaY, newdata = data.frame(A = 1, W),
                        type = "response")

# get estimated propensity for observing outcome if A = 0
ps_DeltaY_A0 <- predict(fit_DeltaY, newdata = data.frame(A = 0, W),
                        type = "response")

# now combine all results into a single propensity score
gn <- list(
  # propensity for DeltaA = 1, A = 1, DeltaY = 1
  ps_DeltaA * ps_A1 * ps_DeltaY_A1,
  # propensity for DeltaA = 1, A = 0, DeltaY = 1
  ps_DeltaA * ps_A0 * ps_DeltaY_A0
)

# pass in this gn to drtmle
out_fit_ps <-  drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                      # make sure a_0/gn are ordered properly!
                      gn = gn, a_0 = c(1, 0),
                      glm_Q = "W1 + W2*A", glm_Qr = "gn",
                      glm_gr = "Qn", family = binomial())
out_fit_ps
```

## Multi-level treatments

```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y2 <- rbinom(N, 1, plogis(W$W1 + 2*W$W2))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
EY2 <- mean(Y2)
```

In our examples so far, we have only considered binary treatments. However,
\texttt{drtmle} is fully equipped to handle treatments with an arbitrary number
of discrete values as well. Suppose that $A$ assumes values in a set
$\mathcal{A}$ of limited cardinality, the PS regression estimation is modified
to ensure that
\begin{equation*}
  \sum\limits_{a \in \mathcal{A}} g_n(a \mid w) = 1 \ \mbox{for all $w$} \ .
\end{equation*}
If this condition is true, we say that the estimates of the PS are compatible.
Many regression methodologies that have been developed work well with binary
outcomes, but have not been extended to multi-level outcomes. To ensure that
users of \texttt{drtmle} have access to the large number of binary regression
techniques when estimating the PS, we have taken an alternative approach to
estimation of the propensity for each level of the treatment. Specifically,
rather than fitting a single multinomial-type regression, we fit a series of
binomial regressions. As a concrete example, consider the case of no missing
data and where $\mathcal{A} = \{0,1,2\}$. We can ensure compatible estimates of
the PS by generating an estimate $g_n(0 \mid w)$ of $Pr_0(A = 0 \mid W = w)$
and $\tilde{g}_n(1 \mid w)$ of $Pr_0(A = 1 \mid A > 0, W = w)$. The latter
estimate may be generated by regressing $I(A = 1)$ on $W$ in observations with
$A > 0$. Since the outcome is binary, this regression may be estimated using
any technique suitable for binary outcomes. By rules of conditional
probability, it holds that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
Pr_0(A = 0 \mid W)\}$ and thus an estimate of $g_0(1 \mid w)$ can be computed
as $g_n(1 \mid w) = \tilde{g}_n(1 \mid w) \{1 - g_n(0 \mid w)\}$. Finally, we
let $g_n(2 \mid w) = 1 - g_n(0 \mid w) - g_n(1 \mid w)$, which ensures
compatibility of the estimates for each level of the covariates. This approach
generalizes to an arbitrary number of treatment levels. Below, we provide an
illustration using simulated data with a treatment that has three levels. The
true parameter values in this simulation are $\psi_0(0)=$ `r round(EY0,2)`,
$\psi_0(1)=$ `r round(EY1, 2)`, and $\psi_0(2)=$ `r round(EY2,2)`.

```{r, cache = TRUE, message = FALSE}
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 2, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
```

The multi-level PS can still be estimated using any of the three techniques
discussed previously. Here, we illustrate with generalized linear
models.

```{r, cache = TRUE, message = FALSE}
glm_fit_multiA <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         family = binomial(), a_0 = c(0,1,2))
glm_fit_multiA
```

Above, we used the option `a_0` to specify the levels of the treatment at which
we were interested in estimating marginal means. There is no need for `a_0` to
contain all unique values of `A`. For example, certain levels of the treatment
may not be of scientific interest, but nevertheless we would like to use these
data to help estimate the OR and PS (by setting `stratify = FALSE`).

The confidence interval and testing procedures extend to multi-level treatments
with minor modifications. We can still obtain a confidence interval and
hypothesis test for each of the marginal means by not specifying a `contrast`.

```{r, cache = TRUE}
ci(glm_fit_multiA)
wald_test(glm_fit_multiA, null = c(0.4, 0.5, 0.6))
```

Alternatively, we can specify a `contrast` to estimate confidence intervals
about the average treatment effect of $A=1$ versus $A=0$. Calls to `wald_test()`
can also be made.

```{r, cache = TRUE}
ci(glm_fit_multiA, contrast = c(-1, 1, 0))
```

Similarly, we can compute confidence intervals comparing the average treatment
effect of $A = 2$ versus $A = 0$.

```{r, cache = TRUE}
ci(glm_fit_multiA, contrast = c(-1, 0, 1))
```

User-specified contrasts are also available. For example, we can modify the
`riskRatio` object we used previously for a two-level treatment to compute the
risk ratio comparing $A=1$ to $A=0$ and comparing $A = 2$ to $A = 0$.

```{r, cache = TRUE}
riskRatio_1v0 <- list(f = function(eff){ log(eff) },
                      f_inv = function(eff){ exp(eff) },
                      h = function(est){ est[2]/est[1] },
                      fh_grad =  function(est){ c(1/est[2], -1/est[1], 0) })
ci(glm_fit_multiA, contrast = riskRatio_1v0)
```

Similarly, we can compute confidence intervals for the risk
ratio comparing the marginal mean for $A=2$ to $A=1$.

```{r, cache = TRUE}
riskRatio_2v0 <- list(f = function(eff){ log(eff) },
                      f_inv = function(eff){ exp(eff) },
                      h = function(est){ est[3]/est[1] },
                      fh_grad =  function(est){ c(0, -1/est[1], 1/est[3]) })
ci(glm_fit_multiA, contrast = riskRatio_2v0)
```

## Cross-validated regression estimates

To avoid overfitting initial regressions, one can use cross-validated estimates
of the nuisance regressions. Theoretically, the use of cross-validated
regression weakens entropy conditions on nuisance parameter estimators
necessary to achieve asymptotic normality [@klaassen1987consistent;
@bickel1993efficient; @zheng:2010:working]. This is true of the regular TML and
AIPTW estimators, as well as of the TML estimator with doubly robust inference.
The `drtmle()` function implements cross-validated estimation of the nuisance
regression parameters through the `cvFolds` option, as shown in the code below.

```{r, cache = TRUE}
cv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_Q = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_gr = c("SL.glm", "SL.mean"),
                    SL_Qr = c("SL.glm", "SL.mean"),
                    stratify = FALSE, cvFolds = 2, a_0 = c(0, 1, 2))
```

Cross-validation may significantly increase the computation time necessary for
running `drtmle()`. Parallelized nuisance regression fitting is available by
setting the option `use_future = TRUE`. Since, internally, all computations are
carried out using futures (with the API provided by the \texttt{future}
\texttt{R} package), setting this option amounts only to parallelizing with
futures [@futurePackage; @bengtsson2021unifying] via `future.apply::lapply()`;
when `use_future = FALSE`, base `lapply()` is simply used instead. Together,
setting these options results in parallelization of the estimation of the OR,
PS, and reduced-dimension regressions.

If no adjustments are made prior to invoking the `drtmle()` function, futures
are computed sequentially, as per the defaults of the \texttt{future} package.
It is left to the user to specify appropriate system-specific future backends
that may better suit both their analytical problem and computing environment
(e.g., using the \texttt{future.batchtools} package [@futurebatchtoolsPackage]
or even simply setting a parallelized `future::plan()`). Below, we demonstrate
registration using an ad-hoc cluster on the local system. The choice of backend
is flexible, and submission of the job to a variety of schedulers (e.g., SLURM,
TORQUE) is also permitted. Interested readers are invited to consult the
documentation of the \texttt{future.batchtools} package for further information.

```{r, message = FALSE, eval = FALSE}
# set up parallel backend via future.batchtools
library(future.batchtools)
library(parallel)
cl <- makeCluster(2, type = "SOCK")
plan(cluster, workers = cl)
clusterEvalQ(cl, library("SuperLearner"))

# run drtmle() in parallel using the registered backend via use_future = TRUE
pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
                     SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
                     SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
                     SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
                     stratify = FALSE, a_0 = c(0,1,2),
                     cvFolds = 2, use_future = TRUE)
```

# Discussion

The \texttt{drtmle} package was designed to provide a user-friendly
implementation of the TMLE algorithm that facilitates doubly robust inference
about marginal means and average treatment effects. While the theory underlying
doubly robust inference is complex, the implementation of the methods only
requires users to provide the data and specify how to estimate nuisance
regressions. Though estimation of the OR and PS is familiar to those with
backgrounds in causal inference-related estimation, the reduced-dimension
regressions may appear strange. Indeed, it is difficult to provide intuition for
these parameters and more difficult still to determine what scientific
background knowledge might guide users in how to estimate these regression
parameters. Therefore, we recommended estimating these quantities adaptively,
e.g., with the super learner algorithm. The super learner library should contain
several relatively smooth estimators such as `SL.mean`, `SL.glm`, and `SL.gam`.
In particular, including `SL.mean` (an unadjusted generalized linear model) may
be important due to the fact that, when the OR and PS are consistently
estimated, the reduced-dimension regressions are identically equal to zero.
Thus, this unadjusted regression estimator may do a good job estimating the
reduced-dimension regressions in situations where the OR and PS are estimated
well. In addition to relatively smooth estimators, we recommend additionally
including several adaptive estimators such as `SL.npreg` or `SL.earth`.

Planned extensions of the \texttt{drtmle} package include extensions to
longitudinal settings involving treatments at multiple time points and inclusion
of the collaborative targeted maximum likelihood estimator (C-TMLE) of
@vdl:2014:ijb that also facilitates collaborative doubly robust inference. The
\texttt{drtmle} \texttt{R} package is publicly available on the Comprehensive
\texttt{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=drtmle,
and the open source development repository is available on the GitHub platform
for collaborative programming at https://github.com/benkeser/drtmle; the
package's documentation (automatically built via the \texttt{pkgdown} \texttt{R}
package) is hosted via GitHub Pages at https://benkeser.github.io/drtmle/.

# Appendix

## Arbitrary user-specified regressions

The \texttt{drtmle} package also allows users to pass in their own estimated OR
and PS via the `Qn` and `gn` options, respectively. If `Qn` is specified by the
user, `drtmle()` will ignore the nuisance parameter estimation routines
specified by `SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As
noted in the package documentation, there is a specific format necessary for
inputting `Qn` and `gn`. For `Qn`, the entries in the list should correspond to
the OR evaluated at `A` and the observed values of `W`, with order determined by
the input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the
OR at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
example below illustrates this concept.

```{r, cache = TRUE, message = FALSE}
# use simplified data
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 1, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))

# fit a GLM for outcome regression outside of drtmle
out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())

# generate Qn list
Qn10 <- list(
  # first entry is predicted values setting A = 1
  predict(out_glm_fit, newdata = data.frame(A = 1, W), type = "response"),
  # second entry is predicted values setting A = 0
  predict(out_glm_fit, newdata = data.frame(A = 0, W), type = "response")
)

# pass this list to drtmle to avoid internal estimation of
# outcome regression (note propensity score and reduced-dimension
# regressions are still estimated internally)
out_fit1 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn", Qn = Qn10,
                  # crucial to set a_0 to match Qn's construction!
                  a_0 = c(1,0))
out_fit1

# to be completely thorough let's re-order Qn10 and re-run
Qn01 <- list(Qn10[[2]], Qn10[[1]])

out_fit2 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn",
                  # use re-ordered Qn list
                  Qn = Qn01,
                  # a_0 has to be reordered to match Qn01!
                  a_0 = c(0,1))
# should be the same as out_fit1, but re-ordered
out_fit2
```

Similarly, for `gn` we need to pass in properly ordered lists. For example, if
`a_0 = c(0, 1)` then `gn[[1]]` should be the PS for receiving $A = 0$ and
`gn[[2]]` should be the PS for receiving $A = 1$. The example below illustrates
this concept.

```{r, cache = TRUE, message = FALSE}
# fit a GLM for propensity score outside of drtmle
out_glm_fit <- glm(A ~ . , data = data.frame(W), family = binomial())

# get P(A = 1 | W) by calling predict
gn1W <- predict(out_glm_fit, newdata = data.frame(W), type = "response")

# generate gn list
gn10 <- list(
  # first entry is P(A = 1 | W)
  gn1W,
  # second entry is P(A = 0 | W) = 1 - P(A = 1 | W)
  1 - gn1W
)

# pass this list to drtmle to avoid internal estimation of
# propensity score (note reduced-dimension regressions are
# still estimated internally)
out_fit3 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                   glm_Qr = "gn", glm_gr = "Qn",
                   Qn = Qn10, gn = gn10,
                   # crucial to set a_0 to match Qn and gn's construction!
                   a_0 = c(1,0))
out_fit3

# to be completely thorough let's re-order gn10 and re-run
gn01 <- list(gn10[[2]], gn10[[1]])

out_fit4 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                   glm_Qr = "gn", glm_gr = "Qn",
                   # use re-ordered Qn/gn list
                   Qn = Qn01, gn = gn01,
                   # a_0 has to be reordered to match Qn01!
                   a_0 = c(0,1))
# should be the same as out_fit3, but re-ordered
out_fit4
```

This flexibility allows users to leverage any regression technique to externally
obtain initial estimates of nuisance parameters. The key step is making sure
that the ordering of `Qn` and `gn` is consistent with `a_0`.

## Additional methods included in \texttt{drtmle}

### Inference for IPTW using adaptive PS estimators

Doubly robust estimators of marginal means are appealing in their robustness and
efficiency properties; however, researchers may prefer the IPTW estimator for
its intuitive derivation and implementation. Heretofore, inference for IPTW
estimators precluded the use of adaptive estimators of the PS, yet @vdl:2014:ijb
proposed modified IPTW estimators that allow for adaptive PS estimation by
incorporating an additional step to update the initial PS estimates. Like those
presented above, this estimator is based on an additional regression parameter,
$\tilde{Q}_{r,0n}(a,w) \coloneqq E_0\{Y \mid A = a, g_n(W) = g_n(w)\}$, that is,
the regression of the outcome $Y$ on the estimated propensity score amongst
observations with $A=a$. The author showed that if $g_n^{\star}(a \mid w)$, an
estimator of the propensity for treatment $A=a$, satisfied that
\begin{equation*}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
    W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,
\end{equation*}
then the IPTW estimator based on $g_n^{\star}$ would asymptotically attain a normal
limiting distribution with estimable variance. In fact, @vdl:2014:ijb proposed a
TMLE-type algorithm that mapped an initial PS estimate into an estimate that
satisfies this key equation and variance estimators that can be used to generate
Wald-style confidence intervals and hypothesis tests.

An estimator with equivalent asymptotic properties based on a PS estimate,
$g_n(a \mid w)$, can be computed as
\begin{equation*}
  \psi_{n,\text{IPTW}}(a) = \psi_{n,\text{IPTW}}(a) - \frac{1}{n}
  \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)}
  \{ I(A_i = a) - g_n(a \mid W = W_i)\} \ .
\end{equation*}
Both this estimator and the TMLE-based IPTW estimator above are implemented in
\texttt{drtmle}. Alternative approaches for developing IPTW estimators that are
asymptotically normal and attain the nonparametric efficiency bound have been
investigated by @ertefaie2022nonparametric and @hejazi2022efficient, who relied
on undersmoothing-based corrections rather than the TMLE-type corrections of
@vdl:2014:ijb.

#### Implementation of inference for adaptive IPTW

The IPTW estimators based on adaptive estimation of the PS are implemented in
the `adaptive_iptw()` function. The `adaptive_iptw()` function shares many
options with `drtmle()`; however, as the IPTW estimator does not rely on an
estimate of the OR, options related to the OR are omitted.

```{r, cache = TRUE}
npreg_iptw_multiA <- adaptive_iptw(W = W, A = A, Y = Y, stratify = FALSE,
                                   SL_g = "SL.npreg", SL_Qr = "SL.npreg",
                                   family = binomial(), a_0 = c(0, 1))
npreg_iptw_multiA
```

The `"adaptive_iptw"` S3 class contains identical methods for computing
confidence intervals and Wald-style tests as the `"drtmle"` class, thus readers
are referred back to the section on confidence intervals and hypothesis tests
for reminders on the various options for these procedures. By default, these
methods are implemented for the `iptw_tmle()` estimator of @vdl:2014:ijb.  We
conjecture that this estimator will have improved finite-sample behavior
relative to `iptw_os()` (the alternative estimator described above), though
further study in numerical experiments is warranted.

As with `drtmle()`, the `adaptive_iptw()` function allows missing data in `A`
and `Y`. Similarly, `adaptive_iptw()` also allows for parallelized,
cross-validated estimation of the regression parameters via the `cvFolds` and
`parallel` options.  As with the estimators implemented by `drtmle()`,
cross-validation allows for valid inference under weaker assumptions.

### Outcome-adaptive propensity score estimation

@ju2020robust have recently proposed estimators of the average treatment effect
to be used in settings where this effect is weakly identifiable, as evidenced by
estimated propensity scores of small magnitude. In these settings, doubly robust
estimators may exhibit non-robust behavior in small samples. The proposed
solution is to target an alternative quantity in the propensity score estimation
step. If we are interested in estimating $\psi_0(a)$ for $a = 0, 1$, we can
estimate $Pr_0(A = a \mid \overline{Q}_0(1, W), \overline{Q}_0(0, W))$, which
describes the conditional probability of receiving treatment $A = a$ given the
value of the two outcome regressions. For this reason, we call this quantity an
*outcome-adaptive propensity score*. We have shown that all of the usual
asymptotics for TMLE carry through when an estimator of this quantity is
substituted for the true propensity score. The resultant TML estimator is
*super-efficient*, meaning that it will generally have greater asymptotic
efficiency than a standard TML estimator (in fact, greater asymptotic efficiency
than is ordinarily available to any regular asymptotically linear estimator).
Simulation experiments indicate that this approach also delivers more stable
behavior in settings where estimated propensity scores become very small.
However, the estimator is not doubly robust -- the asymptotics rely on the
outcome regression being consistently estimated at $n^{-1/4}$ rate. The highly
adaptive lasso estimator [@benkeser2016highly; @vdl2017generally;
@bibaut2019fast], implemented in the open source \texttt{hal9001} \texttt{R}
package [@hejazi2020hal9001; @coyle2021hal9001], can attain such a convergence
rate without sacrificing flexible estimation; nevertheless, as a consequence of
this convergence rate requirement, this estimator trades off robustness against
achieving greater stability and efficiency. See @ju2020robust for further
discussion.

#### Implementation of outcome-adaptive propensity score

If the outcome-adaptive PS is desired, the user may specify this via the
`adapt_g` option. In this case, the additional targeting steps needed for doubly
robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed
above, this estimator is *not* doubly robust. Note that when `adapt_g = TRUE`,
the data used for PS estimation will be a `data.frame` with column names `QaW`
where `a` takes all the values specified in `a_0`. For example if `a_0 =
c(0,1)`, then the PS estimation data frame will have two columns called `Q0W`
and `Q1W`. This may be relevant for users writing their own \texttt{SuperLearner}
wrapper functions or those who make use of the `glm_g` setting.

```{r, message = FALSE}
# outcome adaptive propensity score fit using glms
adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
                     adapt_g = TRUE)

# outcome adaptive propensity score fit using super learner
adaptg_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_Q = c("SL.glm", "SL.mean"),
                     SL_g = c("SL.glm", "SL.mean"),
                     adapt_g = TRUE)
```

# Acknowledgements {-}

NSH grateful acknowledges the support of the National Science Foundation (award
no. [DMS 2102840](https://www.nsf.gov/awardsearch/showAward?AWD_ID=2102840)).

# References